---
title: "2024 School Characteristics"
author: "Janette Avelar"
date: '2024-04-30'
output: 
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warnings = FALSE, message = FALSE)
pacman::p_load(tidyverse, here, rio, DT, ggthemes) #plotly, parameters, ggrepel, 
#read in data
load("data/2024 data/complete_canopy_2024.RData")
```

```{r}
#manual branding - file won't load
transcend_cols = c("#1A4C81","#59C3B4","#EF464B","#ADE0EE")
transcend_cols2 = c("#BC2582","#FFA630","#FFDE42","#99C24D","#218380","#D3B7D7")
transcend_grays = c("#4D4D4F","#9D9FA2","#D1D3D4")
transcend_na = transcend_grays[2]
theme_transcend = theme_gdocs(base_size = 14, base_family = "Open Sans") +
  theme(
    plot.title = element_text(family = "Bebas Neue", color = "black"),
    plot.background = element_blank(),
    axis.text = element_text(colour = "black"),
    axis.title = element_text(colour = "black"),
    panel.border = element_rect(colour = "#4D4D4F"),
    strip.text = element_text(size = rel(0.8)),
    plot.margin = margin(10, 24, 10, 10, "pt")
  )
theme_set(theme_transcend)
```

```{r}
#barplot function
barplot <- function(x, metric, title = "Some title"){
  # Ensure metric is a character string
  if (!is.character(metric) || length(metric) != 1) {
    stop("Metric must be a single character string indicating 'n' or 'pct'.")
  }
  
  data <- variables %>% 
    mutate(rate = 1) %>% 
    group_by({{x}}) %>% 
    summarize(n = sum(rate)) %>% 
    ungroup() %>% 
    mutate(total = sum(n),
           pct = n/total) %>% 
    filter(!is.na({{x}}))
  
  y_var <- ifelse(metric == "n", "n", "pct")
  y_lab <- ifelse(metric == "n", "Number of schools", "Percentage of schools")

  # Setup plot
  plot <- ggplot(data, aes(x = {{x}}, y = .data[[y_var]], fill = {{x}})) +
    geom_col() +
    scale_fill_manual(values = transcend_cols) +
    theme(legend.position = "none", panel.grid.major.x = element_blank()) +
    labs(title = title, x = "", y = y_lab)
  
  # Configure y-axis and labels based on metric
  if (metric == "n") {
    max_y <- max(data[[y_var]], na.rm = TRUE)
    limit_upper <- max_y * 1.1
    plot <- plot + scale_y_continuous(expand = c(0, 0), limits = c(0, limit_upper))
    plot <- plot + geom_text(aes(label = n), nudge_y = 0.5, vjust = 0, color = transcend_na, fontface = "bold", size = 5.5, family = "sans")
  } else if (metric == "pct") {
    plot <- plot + scale_y_continuous(expand = c(0, 0), limits = c(0, 1), labels = scales::percent_format())
    plot <- plot + geom_text(aes(label = scales::label_percent(accuracy = 1)(pct)), nudge_y = 0.01, vjust = 0, color = transcend_na, fontface = "bold", size = 5.5, family = "sans")
  }
  
  return(plot)
}
```


## Locale  

Guiding question: **What's the breakdown of schools in the data by locale (urban, suburban, rural)?**

```{r}
barplot(school_locale, 'n', "Canopy schools by geographic region")
barplot(school_locale, 'pct', "Canopy schools by geographic region")
```


## Level  

Guiding question: **What's the breakdown of schools in the data by level (elementary, middle, high)?**

```{r}

```


## Type  

Guiding question: **What's the breakdown of schools in the data by school type (charter, district, independent)?**

```{r}
barplot(school_type, 'n', "Canopy schools by type")
barplot(school_type, 'pct', "Canopy schools by type")
```


## Descriptor  

Guiding question: **What's the breakdown of schools by descriptor?**

## Demographics  

### Students  

### Leadership  

Guiding question: **What's the racial/ethnic breakdown of Canopy school leadership teams?**

### Region  

Guiding question: **Which states are represented or not?**