---
title: "Pilot Practices w/ Alt Clusters"
author: "Janette Avelar"
date: '2024-08-01'
output: 
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.width = 10, fig.height = 8)
pacman::p_load(tidyverse, here, rio, DT, ggthemes, scales, broom, lme4, wordcloud, tm, mgcv)
#read in data
load(here("data", "2024 data/complete_canopy_2024.RData"))
labels <- import(here("data", "longitudinal/tag-labels.csv"))
```

```{r}
#manual branding - file won't load
transcend_cols = c("#1A4C81","#59C3B4","#EF464B","#ADE0EE")
transcend_cols2 = c("#BC2582","#FFA630","#FFDE42","#99C24D","#218380","#D3B7D7")
transcend_grays = c("#4D4D4F","#9D9FA2","#D1D3D4")
transcend_na = transcend_grays[2]
theme_transcend = theme_gdocs(base_size = 14, base_family = "Open Sans") +
  theme(
    plot.title = element_text(family = "Bebas Neue", color = "black"),
    plot.background = element_blank(),
    axis.text = element_text(colour = "black"),
    axis.title = element_text(colour = "black"),
    panel.border = element_rect(colour = "#4D4D4F"),
    strip.text = element_text(size = rel(0.8)),
    plot.margin = margin(10, 24, 10, 10, "pt")
  )
theme_set(theme_transcend)
```

**What do schools most want to pilot in the next 5 years?**

```{r}
tags %>% 
  select(school_id, starts_with("pilot")) %>% 
  pivot_longer(cols = starts_with("pilot"),
               names_to = "variable",
               values_to = "usage") %>% 
  group_by(variable) %>% 
  summarize(n = sum(usage),
            pct = round(n/189, 2)) %>% 
  mutate(variable = str_replace_all(variable, "pilot", "practices")) %>% 
  left_join(labels, by = "variable") %>% 
  arrange(-n) %>% 
  slice_head(n = 5) %>% 
  ggplot(., aes(pct, reorder(label, pct))) +
  geom_col(fill = transcend_cols[1]) +
  scale_x_continuous(expand = c(0, 0), limits = c(0, .2), labels = scales::percent_format()) +
  scale_y_discrete(labels = wrap_format(25)) +
  geom_text(aes(label = scales::label_percent(accuracy = 1)(pct)), 
            hjust = 1.1,
            vjust = 0, 
            color = "white", 
            fontface = "bold", 
            size = 5.5, 
            family = "sans") +
  theme(panel.grid.major.y = element_blank()) +
  labs(x = "",
       y = "",
       title = str_wrap("Top 5 tags Canopy schools hope to pilot in the next 5 years", 60))
```

**Are there broad patterns in the practices schools hope to implement in the next five years?**  


```{r}
#read in cluster information
cluster <- import(here("data", "mixed-methods-clusters-24.csv")) %>% 
  janitor::clean_names() %>% 
  select(tag, cluster = proposed_cluster_for_preliminary_24_analysis) %>% 
  mutate(cluster = case_when(
    cluster == "Postsecondary" ~ "Postsecondary pathways",
    cluster == "Ed justice" ~ "Educational justice",
    cluster == "Individualized" ~ "Individualized learning",
    TRUE ~ as.character(cluster)
  ))
#link to pilot tags
pilot_clusters <- tags %>% 
  select(school_id, starts_with("pilot")) %>% 
  pivot_longer(cols = starts_with("pilot"),
               names_to = "variable",
               values_to = "usage") %>%
  mutate(variable = str_replace_all(variable, "pilot", "practices")) %>% 
  left_join(labels, by = "variable") %>% 
  select(school_id, tag = label, usage) %>% 
  left_join(cluster, by = "tag")
#cluster totals
tots <- pilot_clusters %>% 
  select(tag, cluster) %>% 
  unique() %>% 
  mutate(rate = 1) %>% 
  group_by(cluster) %>% 
  summarize(total = sum(rate))
#weighted total for each school
pilot_clusters %>% 
  group_by(school_id, cluster) %>% 
  summarize(n = sum(usage)) %>% 
  ungroup() %>% 
  left_join(tots, by = "cluster") %>% 
  mutate(pct = n/total) %>% 
  group_by(cluster) %>% 
  summarize(wt_sum = sum(pct),
            wt_mean = mean(pct),
            mean = mean(n),
            median = median(n)) %>% 
  ggplot(., aes(reorder(cluster, -wt_mean), wt_mean)) +
  geom_col(fill = transcend_cols[1]) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, .1), labels = scales::percent_format()) +
  scale_x_discrete(labels = wrap_format(15)) +
  theme(panel.grid.major.x = element_blank()) +
  labs(y = "Average cluster percentage selected",
       x = "",
       title = str_wrap("Clusters Canopy schools are interested in piloting in the next 5 years", 60))
```


```{r}
pilot_clusters %>% 
  filter(usage == 1) %>% 
  group_by(school_id, cluster) %>% 
  summarize(n = sum(usage)) %>% 
  ungroup() %>% 
  left_join(tots, by = "cluster") %>% 
  mutate(pct = n/total) %>% 
  group_by(cluster) %>% 
  summarize(wt_sum = sum(pct),
            wt_mean = mean(pct),
            mean = mean(n),
            median = median(n)) %>% 
  ggplot(., aes(reorder(cluster, -mean), mean)) +
  geom_col(fill = transcend_cols[1]) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 2)) +
  scale_x_discrete(labels = wrap_format(15)) +
  geom_text(aes(label = round(mean, 2)), 
            vjust = -.2, 
            color = transcend_na, 
            fontface = "bold", 
            size = 5.5, 
            family = "sans") +
  theme(panel.grid.major.x = element_blank()) +
  labs(y = "Mean tags selected",
       x = "",
       title = str_wrap("Clusters Canopy schools are interested in piloting in the next 5 years", 60))
```

Modifying the graph to include the number of tags in each cluster schools already use:  


```{r}
tags <- full %>% 
  select(school_id, starts_with("practices")) %>% 
  pivot_longer(cols = starts_with("practices"),
               names_to = "variable",
               values_to = "usage") %>% 
  left_join(labels, by = "variable") %>% 
  rename("tag" = label) %>% 
  left_join(cluster, by = "tag") %>% 
  group_by(school_id, cluster) %>% 
  summarize(n = sum(usage)) %>% 
  group_by(cluster) %>% 
  summarize(mean = mean(n)) %>% 
  mutate(type = "Practices in use")
pilot_clusters %>% 
  filter(usage == 1) %>% 
  group_by(school_id, cluster) %>% 
  summarize(n = sum(usage)) %>% 
  group_by(cluster) %>% 
  summarize(mean = mean(n)) %>% 
  mutate(type = "Practices to pilot") %>% 
  bind_rows(tags) %>% 
  ggplot(., aes(reorder(cluster, -mean), mean, fill = type)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = transcend_cols) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 15)) +
  scale_x_discrete(labels = wrap_format(15)) +
  geom_text(aes(label = round(mean, 2)), 
            position = position_stack(vjust = .5),
            color = "white", 
            fontface = "bold", 
            size = 5.5, 
            family = "sans") +
  theme(panel.grid.major.x = element_blank(),
        legend.position = c(.8,.9)) +
  labs(y = "Mean tags selected",
       x = "",
       title = str_wrap("Clusters Canopy schools are interested in piloting in the next 5 years", 60),
       fill = "")
```

Alternate version, using `facet` instead of stacking the bars:

```{r}
pilot_clusters %>% 
  filter(usage == 1) %>% 
  group_by(school_id, cluster) %>% 
  summarize(n = sum(usage)) %>% 
  group_by(cluster) %>% 
  summarize(mean = mean(n)) %>% 
  mutate(type = "Avg. number of pilot practices selected") %>% 
  bind_rows(tags) %>% 
  mutate(type = ifelse(type == "Practices in use", "Avg. number of practices already in use", as.character(type))) %>% 
  ggplot(., aes(mean, reorder(cluster, -mean))) +
  geom_bar(stat = "identity", fill = transcend_cols[1]) +
  scale_x_continuous(expand = c(0, 0), limits = c(0, 15)) +
  scale_y_discrete(labels = wrap_format(15)) +
  geom_text(aes(label = round(mean, 2)), 
            vjust = .5,
            hjust = -0.1,
            color = transcend_na, 
            fontface = "bold", 
            size = 5.5, 
            family = "sans") +
  theme(panel.grid.major.y = element_blank()) +
  labs(x = "",
       y = "",
       title = str_wrap("What Canopy schools are interested in piloting in the next 5 years", 66),
       fill = "") +
  facet_wrap(~type)
```


Exploring the graph above once more, using number of schools as the denominator rather than number of tags:  


```{r}
pilot_clusters %>% 
  group_by(school_id, cluster) %>% 
  summarize(n = sum(usage)) %>% 
  mutate(selection = ifelse(n > 0, 1, 0)) %>% 
  group_by(cluster) %>% 
  summarize(pct = sum(selection)/189) %>% 
  ggplot(., aes(reorder(cluster, -pct), pct)) +
  geom_col(fill = transcend_cols[1]) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), labels = scales::percent_format()) +
  scale_x_discrete(labels = wrap_format(15)) +
  theme(panel.grid.major.x = element_blank()) +
  geom_text(aes(label = scales::label_percent(accuracy = 1)(pct)), 
            nudge_y = 0.01, 
            vjust = 0, 
            color = transcend_na, 
            fontface = "bold", 
            size = 5.5, 
            family = "sans") +
  labs(y = "Average cluster percentage selected",
       x = "",
       title = str_wrap("Clusters Canopy schools are interested in piloting in the next 5 years", 60))
```

**When we're dividing pilot tags into clusters, which tags are selected the most?**  


```{r}
pilot_clusters %>% 
  group_by(cluster, tag) %>% 
  summarize(n = sum(usage, na.rm = TRUE)) %>% 
  ungroup() %>% 
  group_by(cluster) %>% 
  arrange(-n) %>% 
  slice_head(n = 5) %>% 
  ggplot(., aes(n, reorder(tag, n))) +
  geom_col(fill = transcend_cols[1]) +
  scale_x_continuous(expand = c(0, 0), limits = c(0, 32)) +
  scale_y_discrete(labels = wrap_format(20)) +
  theme(panel.grid.major.y = element_blank(),
        axis.text.y = element_text(size = 10)) +
  geom_text(aes(label = n), 
            vjust = .5,
            hjust = -0.1,
            color = transcend_na, 
            fontface = "bold", 
            size = 4.5, 
            family = "sans") +
  facet_wrap(~cluster, scales = "free_y", labeller = label_wrap_gen(width = 20)) +
  labs(y = "",
       x = "",
       title = str_wrap("What Canopy schools are most interested in piloting in the next 5 years", 75))
```

**When we're dividing pilot tags into clusters, which tags are selected the least?**  


```{r}
pilot_clusters %>% 
  group_by(cluster, tag) %>% 
  summarize(n = sum(usage, na.rm = TRUE)) %>% 
  ungroup() %>% 
  group_by(cluster) %>% 
  arrange(-n) %>% 
  slice_tail(n = 5) %>% 
  ggplot(., aes(n, reorder(tag, n))) +
  geom_col(fill = transcend_cols[1]) +
  scale_x_continuous(expand = c(0, 0), limits = c(0, 32)) +
  scale_y_discrete(labels = wrap_format(20)) +
  theme(panel.grid.major.y = element_blank(),
        axis.text.y = element_text(size = 10)) +
  geom_text(aes(label = n), 
            vjust = .5,
            hjust = -0.1,
            color = transcend_na, 
            fontface = "bold", 
            size = 4.5, 
            family = "sans") +
  facet_wrap(~cluster, scales = "free_y", labeller = label_wrap_gen(width = 20)) +
  labs(y = "",
       x = "",
       title = str_wrap("What Canopy schools are most interested in piloting in the next 5 years", 75))
```


**To what extent do school characteristics play a role in the tags they want to pilot?**  

Quick defs:  
-Binary high/low = High if more than half the tags in that cluster are in use  
-Tri high/average/low = High if mean+SD; low if mean-SD, average within 1 SD

```{r}
#create cluster score
cluster_score <- full %>% 
  select(school_id, starts_with("practices")) %>% 
  pivot_longer(cols = starts_with("practices"),
                 names_to = "variable",
                 values_to = "usage") %>% 
  left_join(labels, by = "variable") %>% 
  rename("tag" = label) %>% 
  left_join(cluster, by = "tag") %>% 
  group_by(school_id, cluster) %>% 
  summarize(n = sum(usage)) %>% 
  left_join(tots, by = "cluster") %>% 
  mutate(pct = n/total,
         mean = mean(pct),
         sd = sd(pct),
         engagement_bi = case_when(
           pct >= 0.5 ~ "High",
           TRUE ~ "Low"
         ),
         engagement_tri = case_when(
           pct > mean + sd ~ "High",
           pct < mean - sd ~ "Low",
           TRUE ~ "Average"
         )) %>% 
  #add outcome 
  left_join(pilot_clusters, by = c("school_id", "cluster")) %>% 
  group_by(school_id, cluster) %>% 
  mutate(pilot_score = sum(usage),
         pilot_score = ifelse(pilot_score >= 1, 1, 0)) %>% 
  select(-c(usage,tag)) %>% 
  unique()
#prep base data
mod_dat <- full %>% 
    select(school_id, school_locale, school_type, grades_pk, grades_elementary, grades_middle, grades_high, school_enrollment, pct_bipoc, pct_ell, pct_frpl, pct_swd, leadership_diversity) %>% 
    mutate(leadership_diversity = gsub("people", "leaders", leadership_diversity),
           school_locale = factor(school_locale, levels = c("Urban", "Suburban", "Rural", "Multiple")),
           school_type = factor(school_type, levels = c("Public district school", "Public charter school", "Independent (private) school")),
           leadership_diversity = factor(leadership_diversity, levels = c("0 - 24% leaders of color", "25 - 49% leaders of color", "50 - 74% leaders of color", "75 - 100% leaders of color")),
           school_enrollment = as.numeric(scale(school_enrollment, center = TRUE, scale = TRUE))) %>% 
    mutate(across(starts_with("pct"), ~as.numeric(scale(., center = TRUE, scale = TRUE)))) %>% 
  drop_na()
      
# model function
log_model <- function(outcome, data, title){ #outcome needs to be dummy
  #model
  mod <- glm(as.formula(paste(outcome, "~ school_locale + school_type + grades_pk + grades_elementary + grades_middle + grades_high + school_enrollment + pct_bipoc + pct_ell + pct_frpl + pct_swd + leadership_diversity + engagement_tri")),
             family = "binomial",
             data = data)
  # set labels
  cov_labels <- c(
    "school_typeIndependent (private) school" = "Independent (private) school",
    "school_typePublic charter school" = "Public charter school",
    "grades_pk" = "PreK",
    "grades_elementary" = "Elementary",
    "grades_middle" = "Middle",
    "grades_high" = "High",
    "school_enrollment" = "School Enrollment",
    "pct_bipoc" = "% BIPOC students",
    "pct_ell" = "% EL-designated students",
    "pct_frpl" = "% FRPL-eligible",
    "pct_swd" = "% Students with disabilities",
    "school_localeMultiple" = "Multiple locales",
    "school_localeSuburban" = "Suburban",
    "school_localeRural" = "Rural",
    "leadership_diversity25 - 49% leaders of color" = "25-49% leaders of color",
    "leadership_diversity50 - 74% leaders of color" = "50-74% leaders of color",
    "leadership_diversity75 - 100% leaders of color" = "75-100% leaders of color",
    "engagement_triHigh" = "High engagement with cluster",
    "engagement_triAverage" = "Average engagement with cluster",
    "engagement_triLow" = "Low engagement with cluster"
  )
  #plot
  plot <- tidy(mod, effects = "ran_pars", conf.int = TRUE) %>%
  filter(term != "(Intercept)") %>%
  mutate(exp_est = exp(estimate), 
         exp_min = exp(estimate - std.error), 
         exp_max = exp(estimate + std.error)) %>% 
  mutate(term = cov_labels[term]) %>% 
  ggplot(., aes(y = fct_reorder(term, exp_est), x = exp_est)) +
  geom_linerange(aes(xmin = exp_min,
                     xmax = exp_max),
                 color = "blue") +
  geom_point() +
  geom_vline(xintercept = 1) +
  scale_x_continuous(
  trans = "log",
  breaks = c(.0625, .25, .5, 1, 2, 4, 16),
  labels = str_wrap(c("1/16 as likely", "1/4 as likely", "1/2 as likely", "Even", "2x as likely", "4x as likely", "16x as likely"), 10),
  expand = expansion(0, 0)
  ) +
  theme_transcend +
  theme(panel.grid.major.y = element_blank()) +
  labs(
    x = "",
    y = "",
    title = str_wrap(title, 60))
  return(plot)}
```

**Postsecondary pathways**

Starting with Postsecondary pathways to see if we get expected results (high schools more likely to adopt).  


```{r}
dat <- mod_dat %>% 
  left_join(cluster_score, by = "school_id") %>% 
  filter(cluster == "Postsecondary pathways")
log_model("pilot_score", dat, "School characteristics predicting piloting 1 or more tags related to Postsecondary Pathways")
```

**Educational Justice**  


```{r}
dat <- mod_dat %>% 
  left_join(cluster_score, by = "school_id") %>% 
  filter(cluster == "Educational justice")
log_model("pilot_score", dat, "School characteristics predicting piloting 1 or more tags related to Educational Justice")
```

**Deeper Learning**  

```{r}
dat <- mod_dat %>% 
  left_join(cluster_score, by = "school_id") %>% 
  filter(cluster == "Deeper learning")
log_model("pilot_score", dat, "School characteristics predicting piloting 1 or more tags related to Deeper Learning")
```

**Individualized Learning**  

```{r}
dat <- mod_dat %>% 
  left_join(cluster_score, by = "school_id") %>% 
  filter(cluster == "Individualized learning")
log_model("pilot_score", dat, "School characteristics predicting piloting 1 or more tags related to Individualized Learning")
```

**Increasing access & supports**  

```{r}
dat <- mod_dat %>% 
  left_join(cluster_score, by = "school_id") %>% 
  filter(cluster == "Increasing access & supports")
log_model("pilot_score", dat, "School characteristics predicting piloting 1 or more tags related to Increasing access & supports")
```

**Student-driven learning**  

```{r}
dat <- mod_dat %>% 
  left_join(cluster_score, by = "school_id") %>% 
  filter(cluster == "Student-driven learning")
log_model("pilot_score", dat, "School characteristics predicting piloting 1 or more tags related to Student-driven learning")
```

**ATTEMPT 2: USING GAM MODELS**

The set of models below do not assume that the relationship between our school characteristics and tags is linear, allowing for a more complicated relationship between engagement with a cluster and the pilot tags selected (i.e., deals with some of the issues that may come from ceiling effects, or whatnot).  

**Deeper learning**  

```{r}
dat <- mod_dat %>% 
  left_join(cluster_score, by = "school_id") %>% 
  filter(cluster == "Deeper learning")
gam_model <- gam(pilot_score ~ school_locale + school_type + grades_pk + grades_elementary + grades_middle + grades_high + school_enrollment + pct_bipoc + pct_ell + pct_frpl + pct_swd + leadership_diversity + s(pct),
                 data = dat)
#note to self: wrapping in s() stands for smoothing - does not assume linear relationship and instead allows model to take on the shape that makes the most sense
plot(gam_model)
```

**Educational justice**  

```{r}
dat <- mod_dat %>% 
  left_join(cluster_score, by = "school_id") %>% 
  filter(cluster == "Educational justice")
gam_model <- gam(pilot_score ~ school_locale + school_type + grades_pk + grades_elementary + grades_middle + grades_high + school_enrollment + pct_bipoc + pct_ell + pct_frpl + pct_swd + leadership_diversity + s(pct),
                 data = dat)
plot(gam_model)
```

**Individualized learning**  

```{r}
dat <- mod_dat %>% 
  left_join(cluster_score, by = "school_id") %>% 
  filter(cluster == "Individualized learning")
gam_model <- gam(pilot_score ~ school_locale + school_type + grades_pk + grades_elementary + grades_middle + grades_high + school_enrollment + pct_bipoc + pct_ell + pct_frpl + pct_swd + leadership_diversity + s(pct),
                 data = dat)
plot(gam_model)
```

**Postsecondary pathways**  

```{r}
dat <- mod_dat %>% 
  left_join(cluster_score, by = "school_id") %>% 
  filter(cluster == "Postsecondary pathways")
gam_model <- gam(pilot_score ~ school_locale + school_type + grades_pk + grades_elementary + grades_middle + grades_high + school_enrollment + pct_bipoc + pct_ell + pct_frpl + pct_swd + leadership_diversity + s(pct),
                 data = dat)
plot(gam_model)
```

**Increasing access & supports**  

Error: A term has fewer unique covariate combinations than specified maximum degrees of freedom

```{r include = FALSE, eval = FALSE}
dat <- mod_dat %>% 
  left_join(cluster_score, by = "school_id") %>% 
  filter(cluster == "Increasing access & supports")
gam_model <- gam(pilot_score ~ school_locale + school_type + grades_pk + grades_elementary + grades_middle + grades_high + school_enrollment + pct_bipoc + pct_ell + pct_frpl + pct_swd + leadership_diversity + s(pct),
                 data = dat)
plot(gam_model)
```

**Student-driven learning**  

Error: A term has fewer unique covariate combinations than specified maximum degrees of freedom

```{r include = FALSE, eval = FALSE}
dat <- mod_dat %>% 
  left_join(cluster_score, by = "school_id") %>% 
  filter(cluster == "Student-driven learning")
gam_model <- gam(pilot_score ~ school_locale + school_type + grades_pk + grades_elementary + grades_middle + grades_high + school_enrollment + pct_bipoc + pct_ell + pct_frpl + pct_swd + leadership_diversity + s(pct),
                 data = dat)
plot(gam_model)
```

