---
title: "Disaggregated Tag Modeling - Longitudinal"
author: "Anwesha Guha"
date: "`r Sys.Date()`"
output: html_document
---
11:20 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

pacman::p_load(here, purrr, tidyverse, ggcorrplot, proxy, plotly, patchwork, psych, GPArotation, parameters, DT, nFactors, broom, scales, rstanarm, broom.mixed, gmodels, ggthemes, tidytext, rio, fastDummies)

source(here("scripts/branding.R"))
full <- import(here("data/longitudinal", "longitudinal_data.csv"))
tags <- import(here("data/longitudinal", "tags-wide.csv"))

full <- left_join(full, tags, by = join_by(year, school_id))
```

# What are the top tags selected when disaggregated (by locale, level, type, leadership team, etc.)?

## Step 1. Data Cleaning

Create dataset for model.
```{r}
#select relevant variables
mod_prep <- full %>% 
  select(school_id, year, school_type, school_locale, school_enrollment, pct_bipoc, pct_ell, pct_frpl, pct_swd, grades_pk, grades_elementary, grades_middle, grades_high, starts_with("practices")) %>% 
  pivot_longer(cols = c(starts_with("practices")),
               names_to = "tag",
               values_to = "value")

# convert type & locale to dummy columns
mod_prep <- mod_prep %>% 
  dummy_cols(select_columns = c("school_type", "school_locale")) %>% 
  select(-c(school_type, school_type_NA, school_locale, school_locale_NA)) %>% 
  rename(independent_private = `school_type_Independent (private) school`,
         public_charter = `school_type_Public charter school`,
         public_district = `school_type_Public district school`)
```

Identify tags that have not been selected that frequently (less than 20% overall). That results in 36 tags (out of 130).
```{r}
m <- mod_prep %>% 
  group_by(tag, value, school_id) %>% 
  summarise(selected_n = n()) %>% 
  arrange(selected_n) %>% 
  filter(value > 0) %>% 
  select(-2)

tag_counts <- m %>%
  group_by(tag) %>%
  summarise(school_count = n_distinct(school_id))

threshold <- n_distinct(mod_prep$school_id) * 0.2

low_select_tags <- tag_counts %>%
  filter(school_count < threshold) %>%
  pull(tag)
```

Remove those tags from the analysis.
```{r}
mod_prep <- mod_prep %>% 
  filter(!(tag %in% low_select_tags)) %>% 
  select(-tag, -value, tag, value) #arrange so tag and value are at the end

vars <- names(mod_prep)[2:18]

logistic_one_dat = 
  mod_prep %>%
  select(value, tag, one_of(vars)) %>%
  na.omit() %>%
  group_by(tag) %>%
  # scaling non-binary predictors
  mutate_at(vars(
    c(
      "school_enrollment",
      "pct_bipoc",
      "pct_ell",
      "pct_swd",
      "pct_frpl"
    )
  ), scale) %>% 
  mutate(year = as.factor(year))
```

## Step 2. Run models
```{r eval=FALSE, echo=TRUE}
# value ~ year + school_enrollment + pct_bipoc + pct_ell + pct_frpl +  pct_swd + grades_pk + grades_elementary + grades_middle + grades_high + independent_private + public_charter + public_district +  school_locale_Multiple + school_locale_Rural + school_locale_Suburban +   school_locale_Urban

# But actually this formula is not really working because of constant vars


bayes_mods = list()

for(x in unique(mod_prep$tag)) {
  
  data_subset <- filter(logistic_one_dat, tag == x)

constant_vars <- names(data_subset)[sapply(data_subset, function(col) length(unique(col)) == 1)]
  
  if(length(constant_vars) > 0) {
    data_subset <- select(data_subset, -all_of(constant_vars))
  }

new_vars <- setdiff(vars, constant_vars)

one_form = as.formula(paste("value", "~", paste(new_vars, collapse = "+")))
  
  bayes_mods[[x]] = stan_glm(
    one_form,
    data = data_subset,
    family = binomial(link = "logit"),
    prior = student_t( #df and scale ok??
      df = 10,
      location = 0,
      scale = 2.5
    )
  )
  
}

bayes_tidy = lapply(bayes_mods, tidy) %>%
  bind_rows(.id = "response") %>%
  filter(term != "(Intercept)") 
  # mutate(
  #   nice_tag = label_tags()(response),
  #   nice_demog = factor(label_dems(term))
  # )

# save(bayes_tidy, bayes_mods, file = "long_bayesmods.rda")

```


```{r}
load(here("data/longitudinal", "long_bayesmods.rda"))
```

## Step 3. Make Plots.

```{r}
tag_labels <- import(here("data", "tag_labels.csv"))

bayes_facet_dat = bayes_tidy %>% 
  group_by(term) %>%
  arrange(desc(abs(estimate))) %>%
  slice(1:10) %>% 
  left_join(tag_labels, by = c("response" = "var")) 

list_plot <- bayes_facet_dat %>% 
  split(.$term) %>% 
  map(~ggplot(., aes(y = exp(estimate), 
             x = reorder_within(label, estimate, within = term, fun = mean)
             )) +
  geom_col(fill = transcend_cols[1]) + 
  labs(y = "",
       x = "") +
  scale_y_continuous(
    trans = "log",
    breaks = c(.0625, .25, .5, 1, 2, 4, 16),
    labels = c("1/16 as likely", "1/4 as likely", "1/2 as likely", "Even", "2x as likely", "4x as likely", "16x as likely"),
    expand = expansion(0, 0)
  ) +
  tidytext::scale_x_reordered() + 
  geom_hline(yintercept = 1, color = transcend_grays[1]) +
  coord_flip() +
  guides(fill = "none") +
  theme(axis.text = element_text(size = rel(0.7)), strip.text = element_text(size = rel(0.6)),
        panel.grid.major.y = element_blank(),
        axis.text.x = element_text(angle = -45, size = rel(0.6), vjust = 0.5, hjust = 0))
  )
```

Add titles to each of them manually.

### School Level
```{r}
list_plot$grades_pk <- list_plot$grades_pk + labs(title = "Pre-K Schools")
list_plot$grades_elementary <- list_plot$grades_elementary + labs(title = "Elementary Schools")
list_plot$grades_middle <- list_plot$grades_middle + labs(title = "Middle Schools")
list_plot$grades_high <- list_plot$grades_high + labs(title = "High Schools")

library(patchwork)

list_plot$grades_pk + list_plot$grades_elementary + list_plot$grades_middle + list_plot$grades_high + plot_annotation(
  title = 'Top 10 Practices Most/Least Likely to be Selected',
  subtitle = 'Note: Practices used by less than 20% of schools are omitted.',
  theme=theme(plot.title=element_text(hjust=0.5),
              plot.subtitle = element_text(hjust=0.5))
)
```

### School Type

```{r}
list_plot$public_charter <- list_plot$public_charter + labs(title = "Public Charter")  + geom_col(fill = transcend_cols[2])
list_plot$public_district <- list_plot$public_district + labs(title = "Public District") + geom_col(fill = transcend_cols[2]) 
list_plot$independent_private <- list_plot$independent_private + labs(title = "Independent/Private") + geom_col(fill = transcend_cols[2])

list_plot$public_charter + list_plot$public_district + list_plot$independent_private + plot_layout(ncol = 1) + plot_annotation(
  title = 'Top 10 Practices Most/Least Likely to be Selected',
  subtitle = 'Note: Practices used by less than 20% of schools are omitted.',
  theme=theme(plot.title=element_text(hjust=0.5),
              plot.subtitle = element_text(hjust=0.5))
)

```

### School Locale

```{r}
list_plot$school_locale_Rural <- list_plot$school_locale_Rural + labs(title = "Rural")  + geom_col(fill = transcend_cols[3])
list_plot$school_locale_Suburban <- list_plot$school_locale_Suburban + labs(title = "Suburban")  + geom_col(fill = transcend_cols[3])
list_plot$school_locale_Urban <- list_plot$school_locale_Urban + labs(title = "Urban")  + geom_col(fill = transcend_cols[3])
list_plot$school_locale_Multiple <- list_plot$school_locale_Multiple + labs(title = "Multiple")  + geom_col(fill = transcend_cols[3])

list_plot$school_locale_Rural +list_plot$school_locale_Suburban +list_plot$school_locale_Urban + list_plot$school_locale_Multiple + plot_annotation(
  title = 'Top 10 Practices Most/Least Likely to be Selected',
  subtitle = 'Note: Practices used by less than 20% of schools are omitted.',
  theme=theme(plot.title=element_text(hjust=0.5),
              plot.subtitle = element_text(hjust=0.5))
)
```

### Student Demographics

```{r}
list_plot$pct_bipoc <- list_plot$pct_bipoc + labs(title = "Higher-than-average % \nstudents classified as BIPOC") + geom_col(fill = transcend_cols[4])
list_plot$pct_ell <- list_plot$pct_ell + labs(title = "Higher-than-average % \nstudents classified as ELL") + geom_col(fill = transcend_cols[4])
list_plot$pct_frpl <- list_plot$pct_frpl + labs(title = "Higher-than-average % \nstudents classified FRPL") + geom_col(fill = transcend_cols[4])
list_plot$pct_swd <- list_plot$pct_swd + labs(title = "Higher-than-average % \nstudents with disabilities") + geom_col(fill = transcend_cols[4])

list_plot$pct_bipoc + list_plot$pct_ell + list_plot$pct_frpl + list_plot$pct_swd + plot_annotation(
  title = 'Top 10 Practices Most/Least Likely to be Selected',
  subtitle = 'Note: Practices used by less than 20% of schools are omitted.',
  theme=theme(plot.title=element_text(hjust=0.5),
              plot.subtitle = element_text(hjust=0.5))
)


```

### School Year

```{r}
list_plot$year2021 <- list_plot$year2021 + labs(title = "2021") + geom_col(fill = transcend_cols2[5])
list_plot$year2022 <- list_plot$year2022 + labs(title = "2022") + geom_col(fill = transcend_cols2[5])
list_plot$year2023 <- list_plot$year2023 + labs(title = "2023") + geom_col(fill = transcend_cols2[5])
list_plot$year2024 <- list_plot$year2024 + labs(title = "2024") + geom_col(fill = transcend_cols2[5])

list_plot$year2021 + list_plot$year2022 + list_plot$year2023 + list_plot$year2024 + plot_annotation(
  title = 'Top 10 Practices Most/Least Likely to be Selected',
  subtitle = 'Note: Practices used by less than 20% of schools are omitted.',
  theme=theme(plot.title=element_text(hjust=0.5),
              plot.subtitle = element_text(hjust=0.5))
)
```

