---
title: "Practices over Time"
author: "Anwesha Guha"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
pacman::p_load(rio, here, tidyverse, DT, plotly)
```

Also see `core-practices-over-time.html`.

```{r}
dat <- import(here("data/longitudinal", "full-tags-wide.csv"))
dictionary <- import(here("data/2024 data", "dictionary_2024.csv"))
source(here("scripts/branding.R"))
```


## Which core practices have been implemented most over time?

```{r}
core_prac <- dat %>% 
  select(school_id, year, starts_with("core")) %>% 
  mutate(school_id = as.factor(school_id),
         year = as.factor(year)) 

core_prac[is.na(core_prac)] <- 0

core_prac <- core_prac %>% 
#  summarise(across(where(is.numeric), ~ sum(.x, na.rm = TRUE))) %>% 
  pivot_longer(starts_with("core"),
               names_to = "core_practice",
               values_to = "times_selected") 

core_prac_dat <- core_prac %>% 
  group_by(core_practice) %>% 
  summarise(selected = sum(times_selected)) %>% 
              arrange(-selected)
```

First note: there are 26 practices that have never been selected as a core practice. They are the following:
```{r}
core_prac_dat %>% 
  filter(selected == 0) %>% 
  pull(core_practice)
```

These are the rest.

```{r}
datatable(core_prac_dat)
```

Let's look more closely at the top 10 on this list.

```{r}
top_core <- core_prac_dat %>% 
  head(10) %>% 
  pull(core_practice)

top_core_dat <- core_prac %>% 
  filter(core_practice %in% top_core) %>% 
  group_by(core_practice, year) %>% 
  summarise(selected = sum(times_selected))
```

```{r}
top_core_dat %>% 
  filter(year != 2019) %>% 
  ggplot(aes(reorder(core_practice, selected), selected, fill = year)) +
  geom_col() +
  scale_fill_manual(values = transcend_cols) +
  scale_y_continuous(expand = c(0,0)) +
  labs(title = "Core Practices by Year Implemented",
       x = "",
       y = "") +
#  scale_x_discrete(labels = label_tags()) +
  theme(legend.position = "bottom", 
        legend.direction = "horizontal",
        text = element_text(size = 7),
        axis.text.x = element_text(angle = 45, hjust = 1)) + 
  coord_flip() +
  theme(
    panel.grid.major.y = element_blank(),
    legend.position = "bottom"
  ) 
```

It looks like 2021 was really driving the top core practices list across the years.

## Have any lost popularity over the last 5 years? 

Well, this is an interesting question given that 2021 seems like it was the year that schools were more liberal with their core practice selections, so I imagine this affects most practices. But let's look at them below.

```{r}
p <- core_prac %>% 
  group_by(core_practice, year) %>% 
  summarise(selected = sum(times_selected)) %>% 
  filter(year != 2019) %>% 
  mutate(year = as.numeric(year)) %>% 
  ggplot(aes(year, selected, color = core_practice)) +
  geom_point() +
  geom_line() +
  scale_fill_manual(values = transcend_cols2) +
  scale_y_continuous(expand = c(0,0)) +
  labs(title = "Core Practices by Year Implemented",
       x = "",
       y = "") +
#  scale_x_discrete(labels = label_tags()) +
  theme(legend.position = "none", 
        legend.direction = "horizontal",
        text = element_text(size = 7),
        axis.text.x = element_text(angle = 45, hjust = 1)) + 
  theme(
    panel.grid.major.y = element_blank(),
    legend.position = "none"
  ) 

ggplotly(p, tooltip = c("core_practice", "selected"))
```


## Which are the oldest stable core tags and newest growing tags?

```{r}
stable_prac <- core_prac %>%
    group_by(core_practice, year) %>% 
  summarise(selected = sum(times_selected)) %>% 
  filter(year != 2019) %>% 
  group_by(core_practice) %>%
  summarise(
    min_selected = min(selected, na.rm = TRUE),
    max_selected = max(selected, na.rm = TRUE),
    range_selected = max_selected - min_selected
  ) %>%
  filter(range_selected <= 50) %>%
  pull(core_practice) 

p <- core_prac %>% 
  group_by(core_practice, year) %>% 
  summarise(selected = sum(times_selected)) %>% 
  filter(year != 2019) %>% 
  filter(core_practice %in% stable_prac) %>% 
  filter(selected > 20) %>% 
  mutate(year = as.numeric(year)) %>% 
  ggplot(aes(year, selected, color = core_practice)) +
  geom_point() +
  geom_line() +
  scale_fill_manual(values = transcend_cols2) +
  scale_y_continuous(expand = c(0,0)) +
  labs(title = "Stable Tags (Selected >20, varied <50)",
       x = "",
       y = "") +
#  scale_x_discrete(labels = label_tags()) +
  theme(legend.position = "none", 
        legend.direction = "horizontal",
        text = element_text(size = 7),
        axis.text.x = element_text(angle = 45, hjust = 1)) + 
  theme(
    panel.grid.major.y = element_blank(),
    legend.position = "none"
  ) 

ggplotly(p, tooltip = c("core_practice", "selected"))
```

For newest growing tags, in my first pass, I am going to filter the practices to those that increased between 2022 and 2024. I'm omitting 2021 for the filter.

```{r}
increased_prac <- core_prac %>% 
  group_by(core_practice, year) %>% 
  summarise(selected = sum(times_selected)) %>% 
  filter(year %in% c(2022, 2024)) %>% 
  pivot_wider(names_from = year, values_from = selected, names_prefix = "year_") %>%
  mutate(change = year_2024 - year_2022) %>%
  filter(change > 0) %>% 
  arrange(desc(change)) %>% 
  pull(core_practice)

p <- core_prac %>% 
  group_by(core_practice, year) %>% 
  summarise(selected = sum(times_selected)) %>% 
  filter(year != 2019) %>% 
  filter(core_practice %in% increased_prac) %>% 
  mutate(year = as.numeric(year)) %>% 
  ggplot(aes(year, selected, color = core_practice)) +
  geom_point() +
  geom_line() +
  scale_fill_manual(values = transcend_cols2) +
  scale_y_continuous(expand = c(0,0)) +
  labs(title = "Tags Increasing since 2022",
       x = "",
       y = "") +
#  scale_x_discrete(labels = label_tags()) +
  theme(legend.position = "none", 
        legend.direction = "horizontal",
        text = element_text(size = 7),
        axis.text.x = element_text(angle = 45, hjust = 1)) + 
  theme(
    panel.grid.major.y = element_blank(),
    legend.position = "none"
  ) 

ggplotly(p, tooltip = c("core_practice", "selected"))
```

```{r}
core_prac %>% 
  group_by(core_practice, year) %>% 
  summarise(selected = sum(times_selected)) %>% 
  filter(year != 2019) %>% 
  filter(year != 2021) %>% 
  filter(core_practice %in% increased_prac) %>% 
  ggplot(aes(reorder(core_practice, selected), selected, fill = year)) +
  geom_col() +
  scale_fill_manual(values = transcend_cols) +
  scale_y_continuous(expand = c(0,0)) +
  labs(title = "Practices Increasing in Selection Since 2022",
       x = "",
       y = "") +
#  scale_x_discrete(labels = label_tags()) +
  theme(legend.position = "bottom", 
        legend.direction = "horizontal",
        text = element_text(size = 7),
        axis.text.x = element_text(angle = 45, hjust = 1)) + 
  coord_flip() +
  theme(
    panel.grid.major.y = element_blank(),
    legend.position = "bottom"
  )
```

What about those with the largest change?
```{r}
core_prac %>% 
  group_by(core_practice, year) %>% 
  summarise(selected = sum(times_selected)) %>% 
  filter(year %in% c(2022, 2024)) %>% 
  pivot_wider(names_from = year, values_from = selected, names_prefix = "year_") %>%
  mutate(change = year_2024 - year_2022) %>%
  filter(change > 0) %>% 
  arrange(desc(change)) %>% 
  head(10) %>% 
  ggplot(aes(x = year_2022, xend = year_2024, y = reorder(core_practice, change), yend = core_practice)) +
  geom_segment(color = "black", linetype = "dotted") +
  geom_point(aes(x = year_2022), color = "red") +
  geom_point(aes(x = year_2024), color = "blue") +
  guides(col = guide_legend(nrow = 1, title = NULL)) + 
  geom_text(
    aes(x = (year_2022 + year_2024)/2 -1, label = paste("Δ =", year_2024 - year_2022), color = factor(sign(year_2024 - year_2022))),
    nudge_y = .3,
    hjust = 0,
    show.legend = FALSE
  )  +
  labs(
    y = "Core Practice",
    x = "Times Selected",
    title = "Core Practices with largest increase \nfrom 2022 to 2024 Across Schools"
  ) +
  theme(
    panel.grid.major.y = element_blank(),
    legend.position = "bottom",
    axis.text.y = element_text(size = rel(0.6))
  )
```


## What is the relationship between core practices most implemented over time and practices on the horizon? (i.e., are we seeing a lot of "brand new" practices piloted, are schools more or less trying out "established" practices, or both?)

```{r}
core_prac <- dat %>% 
  select(school_id, year, starts_with("core"))

long_dat <- import(here("data/longitudinal", "longitudinal_data.csv"))
```

Hm, I don't see pilot in either dataset. Will explore later
