---
title: "Report Odds and Ends"
author: "Janette Avelar"
date: '2024-09-13'
output: 
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.width = 7, fig.height = 6)
pacman::p_load(tidyverse, here, rio, ggthemes, scales, DT)
#read in data
load(here("data", "2024 data/complete_canopy_2024.RData"))
dat <- variables
rm(dictionary, full, tags, variables)
```

```{r}
#manual branding - file won't load
transcend_cols = c("#1A4C81","#59C3B4","#EF464B","#ADE0EE")
transcend_cols2 = c("#BC2582","#FFA630","#FFDE42","#99C24D","#218380","#D3B7D7")
transcend_grays = c("#4D4D4F","#9D9FA2","#D1D3D4")
transcend_na = transcend_grays[2]
theme_transcend = theme_gdocs(base_size = 14, base_family = "Open Sans") +
  theme(
    plot.title = element_text(family = "Bebas Neue", color = "black"),
    plot.background = element_blank(),
    axis.text = element_text(colour = "black"),
    axis.title = element_text(colour = "black"),
    panel.border = element_rect(colour = "#4D4D4F"),
    strip.text = element_text(size = rel(0.8)),
    plot.margin = margin(10, 24, 10, 10, "pt")
  )
theme_set(theme_transcend)
```

## Future concern plot

```{r, fig.width = 8, fig.height = 4}
dat %>% 
  select(school_id, future_resources, leadership_diversity) %>% 
  filter(future_resources != "Prefer not to say") %>% 
  mutate(future_resources = factor(future_resources,
                                   levels = c("No, not at all", "No, not very", "Neutral", "Yes, somewhat", "Yes, extremely")),
         rate = 1,
         leadership_diversity = str_replace_all(leadership_diversity, "people", "leaders")) %>% 
  group_by(future_resources, leadership_diversity) %>% 
  summarize(n = sum(rate)) %>% 
  ungroup() %>% 
  group_by(leadership_diversity) %>% 
  mutate(sum = sum(n),
         pct = n/sum) %>% 
  ungroup() %>% 
  ggplot(., aes(pct, leadership_diversity, fill = future_resources)) +
  geom_col() +
  scale_fill_manual(values = c(
      "Yes, extremely" = transcend_cols[3], 
      "Yes, somewhat" = transcend_cols2[2], 
      "Neutral" = transcend_na,
      "No, not very" = transcend_cols[2], 
      "No, not at all" = transcend_cols[1])) +
  scale_x_continuous(expand = c(0, 0), limits = c(0, 1), labels = scales::percent_format()) +
  #scale_y_discrete(labels = function(x) str_wrap(x, width = 15)) +
  geom_text(aes(label = scales::label_percent(accuracy = 1)(pct)), 
            position = position_stack(vjust = .5), #vjust = .95 before
            #hjust = 1,
            color = "white", 
            size = 4, 
            family = "sans") +
  theme(legend.position = "bottom", 
        legend.direction = "horizontal", 
        panel.grid.major.y = element_blank(), 
        axis.text.y = element_text(size = 11),
        legend.text = element_text(size = 8),
        legend.title = element_text(size = 10)) +
  labs(x = "", 
       y = "", 
       title = str_wrap("Concern with future sustainability by level of racial diversity in school leadership", 43), 
       fill = "Level of concern") +
  guides(fill = guide_legend(reverse = T, 
                             #ncol = 3,
                             title.position = "top",
                             title.hjust = 0.5))
```

## Static cluster plots

```{r}
#read in data
dat <- import(here("data/longitudinal", "tags-long.csv"))
#read in tag labels
labels <- import(here("data", "tag_labels.csv"))
#read in cluster information
cluster <- import(here("data/longitudinal", "tag_clusters_longitudinal.csv")) %>% 
  janitor::clean_names()
```

```{r}
# start by creating cluster subsets
#subset values
deeper <- cluster %>% 
  filter(proposed_cluster_for_preliminary_24_analysis == "Deeper learning") %>% 
  select(label = tag) %>% 
  left_join(labels, by = "label") %>% 
  pull(var)
ed_justice <- cluster %>% 
  filter(proposed_cluster_for_preliminary_24_analysis == "Ed justice") %>% 
  select(label = tag) %>% 
  left_join(labels, by = "label") %>% 
  pull(var)
individualized <- cluster %>% 
  filter(proposed_cluster_for_preliminary_24_analysis == "Individualized") %>% 
  select(label = tag) %>% 
  left_join(labels, by = "label") %>% 
  pull(var)
postsecondary <- cluster %>% 
  filter(proposed_cluster_for_preliminary_24_analysis == "Postsecondary") %>% 
  select(label = tag) %>% 
  left_join(labels, by = "label") %>% 
  pull(var)
blended <- c("practices_blended_learning", "practices_a_la_carte", "practices_flipped_classroom", "practices_flex", "practices_enriched_virtual", "practices_station_rotation")
```

```{r}
#set up repeating schools - 3+ years
repeaters_3 <- dat %>% 
  select(school_id, year) %>% 
  unique() %>% 
  mutate(rate = 1) %>% 
  pivot_wider(names_from = "year",
              values_from = "rate") %>% 
  rowwise() %>% 
  mutate(participate = sum(`2019`, `2021`, `2022`, `2023`, `2024`, na.rm = TRUE)) %>% 
  filter(participate > 2) %>% 
  pull(school_id)
#set up totals
tots_3 <- dat %>% 
  filter(school_id %in% repeaters_3) %>% 
  group_by(year) %>% 
  summarize(total = n_distinct(school_id))
#function to prep data - generates overall percent for sample of selected schools
prep_overall <- function(data, cluster, group, total){
  dat <- data %>% 
    filter(var %in% cluster) %>% 
    filter(school_id %in% group) %>% 
    group_by(var, year) %>% 
    summarize(n = sum(usage, na.rm = TRUE)) %>% 
    ungroup() %>% 
    left_join(total, by = "year") %>% 
    mutate(pct = n/total) %>% 
    left_join(labels, by = "var") %>% 
    group_by(var) %>% 
    mutate(x_coord = max(year),
           y_coord = pct[which.max(year)])
  
  condition <- dat %>% 
    select(var, year, pct) %>% 
    pivot_wider(names_from = year,
                values_from = pct) %>% 
    group_by(var) %>% 
    mutate(condition = ifelse(`2024` > `2021`, "growing", "shrinking")) %>% 
    ungroup() %>% 
    select(var, condition) %>% 
    unique()
  
  dat <- dat %>% 
    left_join(condition, by = "var")

    return(dat)
}
#function for plotting
cluster_plot_within <- function(data, title) { 
  plot <- ggplot(data, aes_string(x = "year", y = "pct"), color = condition) +
    geom_line(aes(group = label), color = transcend_cols[1], alpha = 0.5) +
    geom_point(aes(group = label), color = transcend_cols[1], alpha = 0.5, size = 1) + 
    geom_smooth(se = FALSE, method = "loess", color = transcend_cols[3], aes(group = 1)) + 
    scale_x_continuous(limits = c(2019, 2024),
                       expand = c(0.01, 0),
                       breaks = unique(data$year)) +
    scale_y_continuous(limits = c(0, 1),
                       expand = c(0, 0),
                       labels = scales::percent_format(accuracy = 1)) +
    theme(panel.grid.major.y = element_blank(), panel.grid.minor.y = element_blank(), legend.position = "none") +
      labs(x = "",
           y = "Percentage of schools selecting tag",
           title = str_wrap(title, 60),
           subtitle = str_wrap("Each blue line represents the trajectory for an individual practice while the red line represents the overall trend. Analysis includes schools with 3 or more survey year responses (N = 119)", 60))

  return(plot)
}
```

### Deeper learning

```{r}
deep <- prep_overall(dat, deeper, repeaters_3, tots_3)
cluster_plot_within(deep, "Changes in deeper learning cluster 2019-2024")
```

### Educational justice

```{r}
justice <- prep_overall(dat, ed_justice, repeaters_3, tots_3)
cluster_plot_within(justice, "Changes in educational justice cluster 2019-2024")
```

### Individualized learning

```{r}
individualized <- prep_overall(dat, individualized, repeaters_3, tots_3)
cluster_plot_within(individualized, "Changes in individualized learning cluster 2019-2024")
```

### Postsecondary pathways

```{r}
postsecondary <- prep_overall(dat, postsecondary, repeaters_3, tots_3)
cluster_plot_within(postsecondary, "Changes in postsecondary pathways cluster 2019-2024")
```

### Blended learning

These practices are also represented in the figures above.

```{r}
blended <- prep_overall(dat, blended, repeaters_3, tots_3)
cluster_plot_within(blended, "Changes in blended learning 2019-2024")
```

```{r}
#clean up
rm(blended, cluster, deep, individualized, justice, labels, postsecondary, tots_3, deeper, ed_justice, repeaters_3, cluster_plot_within, prep_overall)
```

## Waffling

```{r}
#read in data
long = import(here("data/longitudinal", "tags-long.csv"))
#get vector of 4-year tags
tags_4y = long |>
  filter(usage == 1) |>
  filter(n_distinct(year) >= 4, .by = var) |>
  pull(var) |>
  unique()
long_4y = long |> filter(var %in% tags_4y)
#get schools
sch = import(here("data/long_school.csv")) |>
  summarize(n_years = n_distinct(year), .by = c(school_id, school_name))
## append 2024 schools - G's temp fix
sch = long |>
  summarize(
    n_years = n_distinct(year),
    .by = school_id
  ) |>
  left_join(
    select(sch, school_id, school_name)
  )
```


```{r}
potential_wafflers = long_4y |>
  distinct(school_id, year) |>
  filter(n_distinct(year) >= 3, .by = school_id) |>
  summarize(
    years = toString(year),
    could_covid_waffle = 2019 %in% year & 2021 %in% year,
    could_waffle_2022 = 2021 %in% year & 2022 %in% year,
    could_waffle_2023 = 2022 %in% year & 2023 %in% year,
    could_waffle_2024 = 2023 %in% year & 2024 %in% year,
    .by = school_id
  )

waffle_denom = potential_wafflers |>
  summarize(across(starts_with("could"), sum))

waffle_all = long_4y |>
  filter(
    1 %in% usage,    ## present at least once
    n() > 2,         ## at least 3 survey responses
    .by = c(var, school_id)
  ) |> 
  mutate(
    added = any(diff(usage) == 1),
    dropped = any(diff(usage) == -1),
    all_1 = all(usage == 1),
    waffled = added & dropped,
    waffle_bump = waffled & last(usage, na_rm = TRUE) == 0,
    waffle_dip = waffled & last(usage, na_rm = TRUE) == 1,
    .by = c(var, school_id)
  ) |>
  pivot_wider(names_from = year, values_from = usage) |>
  left_join(sch, by = "school_id") |>
  select(school_id, school_name, tag = var, `2019`:`2024`, added, dropped, all_1, starts_with("waffle")) |>
  arrange(school_id, desc(waffled)) |>
  mutate(
    binary_summary = paste(`2019`, `2021`, `2022`, `2023`, `2024`, sep = ","),
    waffle_type = case_when(
      `2019` == 0 & `2021` == 1 & waffled ~ "COVID bump",
      `2019` == 1 & `2021` == 0 & waffled ~ "COVID dip",
      `2021` == 0 & `2022` == 1 & waffled ~ "2022 bump",
      `2021` == 1 & `2022` == 0 & waffled ~ "2022 dip",
      `2022` == 0 & `2023` == 1 & waffled ~ "2023 bump",
      `2022` == 1 & `2023` == 0 & waffled ~ "2023 dip",
      `2023` == 0 & `2024` == 1 & waffled ~ "2024 bump",
      `2023` == 1 & `2024` == 0 & waffled ~ "2024 dip"
    )
  ) 

plot_dat <- waffle_all |>
  filter(waffled) |>
  count(waffle_type) |>
  filter(!is.na(waffle_type)) |>
  separate_wider_delim(waffle_type, delim = " ", names = c("pandemic", "type")) |>
  pivot_wider(names_from = type, values_from = n) |>
  arrange(pandemic) |> ### COVID on top
  mutate(
    all_waffles = bump + dip,
    n_schools_that_could_waffle = c(waffle_denom$could_waffle_2022, waffle_denom$could_waffle_2023, waffle_denom$could_waffle_2024, waffle_denom$could_covid_waffle),
    waffles_per_school = all_waffles / n_schools_that_could_waffle
  )
datatable(plot_dat,
    rownames = FALSE,
    options = list(autoWidth = TRUE, iDisplayLength = 5),
    caption = "COVID vs wafflers in years following"
  ) |>
  formatRound("waffles_per_school", digits = 1)
```

```{r}
#plot
plot_dat %>% 
  select(pandemic, waffles_per_school) %>% 
  mutate(pandemic = factor(pandemic, 
                           levels = c("COVID", "2022", "2023", "2024"),
                           labels = c("2021", "2022", "2023", "2024"))) %>% 
  ggplot(., aes(pandemic, waffles_per_school)) +
  geom_segment(aes(x = pandemic,
                   xend = pandemic,
                   y = 0, 
                   yend = waffles_per_school), 
                   color = transcend_cols[1]) +
  geom_point(size = 5, color = transcend_cols[3]) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 5)) +
  labs(x = "", 
       y = "Average waffles per school", 
       title = "Average waffles per school from 2021-2024") +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank())
```

### Waffling Redux

```{r waffle_redux}
wafflers = long_4y |>
  ## keep only school/tag combos with at least 3 years
  filter(n_distinct(year) >= 3, .by = c(school_id, var)) |>
  arrange(school_id, var, year) |> 
  mutate(
    diff = c(NA, diff(usage)),
    .by = c(school_id, var)
  ) |>
  summarize(
    usage_string = paste(usage, collapse = ""),
    diff_string = paste(diff[-1], collapse = ""),
    year_string = paste(year, collapse = ", "),
    is_waffle = -1 %in% diff & 1 %in% diff,
    is_dip = is_waffle & which.min(diff) < which.max(diff),
    is_bump = is_waffle & which.max(diff) < which.min(diff),
    dip_years = if(is_dip) list(year[usage == 0]) else list(NULL),
    bump_years = if(is_bump) list(year[usage == 1]) else list(NULL),
    waffle_years = coalesce(dip_years, bump_years),
    potential_years = list(tail(head(year, -1), -1)),
    is_multi_waffle = sum(diff == 1, na.rm = TRUE) > 1 | sum(diff == -1, na.rm = TRUE) > 1,
    .by = c(school_id, var)
  )

wafflers |>
  count(is_waffle, is_multi_waffle) |>
  knitr::kable()
```

Here are the details on the multi-waffles. I have omitted them from the waffle
tables that follow:
  
```{r}
wafflers |>
  filter(is_multi_waffle) |>
  datatable()
```

Now omitting the multi-waffles, here are some stats on the rest:
  
```{r}
waffle_summary = wafflers |>
  filter(!is_multi_waffle) |>
  unnest(potential_years) |>
  mutate(waffle_this_year = potential_years %in% waffle_years, 
         .by = c(school_id, var, potential_years)) |>
  summarize(
    waffle_possible = n(),
    waffle_count = sum(waffle_this_year),
    bump_count = sum(waffle_this_year & is_bump),
    dip_count = sum(waffle_this_year & is_dip),
    waffle_pct = mean(waffle_this_year),
    n_sch = n_distinct(school_id),
    waffle_per_sch = waffle_count / n_sch,
    .by = potential_years
  )

waffle_summary |>
  datatable(
    rownames = FALSE,
    options = list(autoWidth = TRUE, iDisplayLength = 5),
    caption = "COVID vs wafflers in years following"
  ) |>
  formatPercentage("waffle_pct", digits = 2) |>
  formatRound("waffle_per_sch", digits = 1)


ggplot(waffle_summary, aes(factor(potential_years), waffle_per_sch)) +
  geom_segment(aes(
                   xend = factor(potential_years),
                   y = 0, 
                   yend = waffle_per_sch), 
                   color = transcend_cols[1]) +
  geom_point(size = 5, color = transcend_cols[3]) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 4)) +
  labs(x = "", 
       y = "Average waffles per school", 
       title = "Average waffles per school from 2021-2023") +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank())
```


