---
title: "Preliminary Investigation"
author: "Janette Avelar"
date: '2023-09-22'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(rio)
dat_19 <- import(here("data", "canopy-2019-2021.rds"))
load("data/complete_canopy_2022.RData")
dat_22 <- clean_data_full
load("data/complete_canopy_2023.RData")
dat_23 <- full
```

# Deep dive into high-participation schools

For the following analysis, I identified the schools across our dataset that participated all 4 years. I've created a new dataset comprised solely of tag selection, and will investigate these.

```{r merge school participation data}
#ID which schools participated all 4 years
#Export ID list to merge with tags

#step 1 - create 0/1 participation for 2019 & 2021
schools_19 <- dat_19 %>% 
  select(year, school_id) %>% 
  mutate(rate = rep(1, nrow(.))) %>% 
  pivot_wider(names_from = year,
              values_from = rate) %>% 
  mutate(`2021` = case_when(
    `2020` == 1 ~ 1,
    `2021` == 1 ~ 1,
    TRUE ~ 0
  )) %>% 
  select(!`2020`)
#step 2 - create 0/1 participation for 2022
schools_22 <- dat_22 %>% 
  select(school_id) %>%
  mutate(school_id = as.character(school_id)) %>% 
  mutate(`2022` = rep(1, nrow(.)))
#step 3 - create 0/1 participation for 2023
schools_23 <- dat_23 %>% 
  select(school_id) %>% 
  mutate(school_id = as.character(school_id)) %>% 
  mutate(`2023` = rep(1, nrow(.)))
#merge
schools <- schools_19 %>% 
  full_join(schools_22, by = "school_id") %>% 
  full_join(schools_23, by = "school_id") %>% 
  mutate(across(where(is.numeric), coalesce, 0))
export(schools, "data/school_participation.csv")
```

Looks like there are 492 distinct schools. Previous analysis had 542 - I think it was not de-duped?  
Check numbers:
*173 schools in 2019 (confirmed)  
*232 schools in 2021 (confirmed - pulled full list [331] & de-duped)  
*161 schools in 2022 (confirmed)  
*251 schools in 2023 (confirmed)  

817 obs over 492 distinct schools = At least 325 of these observations are repeat obs. A total of 29 schools participated all 4 years. (This is higher than what I calculated before - I found the error is coming from inconsistencies in school name generating higher counts. For example, the school ID will stay the same but the school name may be capitalized one year, and not the next.)

```{r ID high participation schools}
dat <- import(here("data", "school_participation.csv")) %>% 
  rowwise() %>% 
  mutate(ttl_part = sum(c(`2019`, `2021`, `2022`, `2023`))
  ) %>% 
  filter(ttl_part == 4) %>% 
  select(school_id)
```

Merge tag data with participation

```{r high participation tag merge}

```

N's years of participation

Looks like 210/492 schools have at least 2 data points - this is about 40% of our data. I'm not sure I understand the question about "how many schools 
*could* we have." We could have all 492 schools with full participation? It depends entirely on who ends up responding, which makes *me* think that we should prioritize the half of our sample that we only have 1 data point for, but others may think otherwise.

```{r summarized years of participation}
import(here("data", "school_participation.csv")) %>% 
  rowwise() %>% 
  mutate(ttl_part = sum(c(`2019`, `2021`, `2022`, `2023`))) %>% 
  ungroup() %>% 
  mutate(rate = rep(1, nrow(.))) %>% 
  group_by(ttl_part) %>% 
  summarize(n = sum(rate, na.rm = TRUE))
```

