---
title: "Preliminary Investigation"
author: "Janette Avelar"
date: '2023-09-22'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(rio)
dat_19 <- import(here("data", "canopy-2019-2021.rds"))
load("data/complete_canopy_2022.RData")
dat_22 <- clean_data_full
load("data/complete_canopy_2023.RData")
dat_23 <- full
```

# Deep dive into high-participation schools

For the following analysis, I identified the schools across our dataset that participated all 4 years. I've created a new dataset comprised solely of tag selection, and will investigate these.

```{r merge school participation data}
#ID which schools participated all 4 years
#Export ID list to merge with tags

#step 1 - create 0/1 participation for 2019 & 2021
schools_19 <- dat_19 %>% 
  select(year, school_id) %>% 
  mutate(rate = rep(1, nrow(.))) %>% 
  pivot_wider(names_from = year,
              values_from = rate) %>% 
  mutate(`2021` = case_when(
    `2020` == 1 ~ 1,
    `2021` == 1 ~ 1,
    TRUE ~ 0
  )) %>% 
  select(!`2020`)
#step 2 - create 0/1 participation for 2022
schools_22 <- dat_22 %>% 
  select(school_id) %>%
  mutate(school_id = as.character(school_id)) %>% 
  mutate(`2022` = rep(1, nrow(.)))
#step 3 - create 0/1 participation for 2023
schools_23 <- dat_23 %>% 
  select(school_id) %>% 
  mutate(school_id = as.character(school_id)) %>% 
  mutate(`2023` = rep(1, nrow(.)))
#merge
schools <- schools_19 %>% 
  full_join(schools_22, by = "school_id") %>% 
  full_join(schools_23, by = "school_id") %>% 
  mutate(across(where(is.numeric), coalesce, 0))
export(schools, "data/school_participation.csv")
```

Looks like there are 492 distinct schools. Previous analysis had 542 - I think it was not de-duped?  
Check numbers:
*173 schools in 2019 (confirmed)  
*232 schools in 2021 (confirmed - pulled full list [331] & de-duped)  
*161 schools in 2022 (confirmed)  
*251 schools in 2023 (confirmed)  

817 obs over 492 distinct schools = At least 325 of these observations are repeat obs. A total of 29 schools participated all 4 years. (This is higher than what I calculated before - again, I think I must have made some de-duping error in calculating or pivoting.)

```{r ID high participation schools}
dat <- import(here("data", "school_participation.csv")) %>% 
  rowwise() %>% 
  mutate(ttl_part = sum(c(`2019`, `2021`, `2022`, `2023`))
  ) %>% 
  filter(ttl_part == 4) %>% 
  select(school_id)
```

Merge tag data with participation

```{r high participation tag merge}

```

