---
title: "Tag data cleanup & merging"
author: "Anwesha Guha"
date: "2023-10-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(tidyverse, here, rio, janitor)
```

Read in data
```{r}
tags <- import(here("data", "tags_2018-2023.csv")) %>% 
  clean_names()
```

```{r}
tags <- tags %>% 
  select(-tag_variable_name_2023_airtable_inconsistencies, -fixed, -priority, -interest) %>% 
  rename(description_2023 = x2023_description,
         description_2022 = x2022_description,
         description_2021 = x2021_description,
         description_2019 = x2019_description,
         equivalent_variable_name_2023 = tag_variable_name_2023,
         current_variable_name = variable_name,
         years_used = years_used_airtable_updated) %>% 
  mutate(most_recent_description = ifelse(!is.na(description_2023), description_2023,
                                          ifelse(!is.na(description_2022), description_2022,
                                                 ifelse(!is.na(description_2021), description_2021, description_2019)))) %>%
  head(130) %>% 
  select(-contains("nested")) %>% 
  relocate(current_variable_name, full_tag_name, most_recent_description)
```

```{r}
tags$current_variable_name <- gsub("IF.*", "", tags$current_variable_name ) #remove the IF RETAINED commentary
```


```{r}
solo_tags <- tags %>% 
  filter(only_one_year_used == "Y") %>% 
  select(current_variable_name, full_tag_name, most_recent_description, years_used)
```


```{r}
removed_tags <- tags %>% 
  filter(year_removed != "Not removed") %>% 
  select(current_variable_name, full_tag_name, equivalent_variable_name_2022:equivalent_variable_name_2019, year_removed, years_used, only_one_year_used)
```


Make tags long.
```{r}
tags_long1 <- tags %>% 
  pivot_longer(description_2023:description_2019, names_to = "year",
               names_prefix = "description_", values_to = "description") %>% 
  select(-starts_with("equivalent_variable_name"), -change_notes, -most_recent_description)

tags_long2 <- tags %>% 
  pivot_longer(equivalent_variable_name_2023:equivalent_variable_name_2019, names_to = "year",
               names_prefix = "equivalent_variable_name_", values_to = "variable_name") %>% 
  relocate(current_variable_name, full_tag_name, most_recent_description, year, variable_name) %>% 
  select(-contains("description"))

tags_long <- full_join(tags_long1, tags_long2)
```

Make tags wide.


```{r}
#save(tags, tags_long, solo_tags, removed_tags, file = "longitudinal_tags.RData")
#save(solo_tags, file = here::here("data", "solo_tags.csv"))
```


Actually, first going to pull in datasets from Drive. Start with 2019-2021.

For 2019:
```{r}
# create update dictionary
tags_19_temp <- tags %>% 
  select(current_variable_name, equivalent_variable_name_2019) %>% 
  drop_na()

# import 2019 dataset
dat_19 <- import(here("data/datasets_by_year", "Schools 2019 (Public).xlsx - Confirmed Schools.csv")) %>% 
  clean_names() %>%
  mutate(across(everything(), as.character)) %>% 
  select(school_id, 15:102) %>% 
  pivot_longer(-school_id, names_to = "tag", values_to = "usage")

# update 2019 dataset with current tag names
dat_19_current <- left_join(dat_19, tags_19_temp, by = join_by(tag == equivalent_variable_name_2019)) %>% 
  select(-tag) %>% 
  rename(var = current_variable_name) %>% 
  mutate(year = "2019")
```

For 2021:
```{r}
# create update dictionary
tags_21_temp <- tags %>% 
  select(current_variable_name, equivalent_variable_name_2021) %>% 
  drop_na()

# import 2021 dataset
dat_21a <- import(here("data/datasets_by_year", "Schools 2021a (Public).xlsx - Fall 2020 School Data.csv")) %>% 
  clean_names() %>%
  mutate(across(everything(), as.character))

dat_21b <- import(here("data/datasets_by_year", "Schools 2021b (Public).xlsx - Winter 2020-21 School Data.csv")) %>% 
  clean_names() %>%
  mutate(across(everything(), as.character))

dat_21 <- bind_rows(dat_21a, dat_21b) %>% 
  select(school_id, 16:106) %>% 
  pivot_longer(-school_id, names_to = "tag", values_to = "usage")

# update 2021 dataset with current tag names
dat_21_current <- left_join(dat_21, tags_21_temp, by = join_by(tag == equivalent_variable_name_2021)) %>% 
  select(-tag) %>% 
  rename(var = current_variable_name) %>% 
  mutate(year = "2021")
```

For 2022:
```{r}
# create update dictionary
tags_22_temp <- tags %>% 
  select(current_variable_name, equivalent_variable_name_2022) %>% 
  drop_na()

# import 2022 dataset
dat_22 <- import(here("data/datasets_by_year", "Schools 2022 (Public).xlsx - Spring 2022 School Data.csv")) %>% 
  clean_names() %>%
  mutate(across(everything(), as.character)) %>% 
  select(school_id, 22:93) %>% 
  pivot_longer(-school_id, names_to = "tag", values_to = "usage")

# update 2022 dataset with current tag names
dat_22_current <- left_join(dat_22, tags_22_temp, by = join_by(tag == equivalent_variable_name_2022)) %>% 
  select(-tag) %>% 
  rename(var = current_variable_name) %>% 
  mutate(year = "2022")
```

For 2023:
```{r}
# create update dictionary
tags_23_temp <- tags %>% 
  select(current_variable_name, equivalent_variable_name_2023) %>% 
  drop_na()

# import 2023 dataset
dat_23 <- import(here("data/datasets_by_year", "Schools 2023 (Public).xlsx - Spring 2023 School Data.csv")) %>% 
  clean_names() %>%
  mutate(across(everything(), as.character)) %>% 
  select(school_id, 39:111) %>% 
  pivot_longer(-school_id, names_to = "tag", values_to = "usage")

# update 2023 dataset with current tag names
dat_23_current <- left_join(dat_23, tags_23_temp, by = join_by(tag == equivalent_variable_name_2023)) %>% 
  select(-tag) %>% 
  rename(var = current_variable_name) %>% 
  mutate(year = "2023")
```

Loooooong dataset commence!
```{r}
longitudinal_dat <- bind_rows(dat_19_current, dat_21_current, dat_22_current, dat_23_current) %>% 
  drop_na(usage, var) %>% 
  filter(str_detect(usage, "1|3-4 years|Less than a year|5+ years|1-2 years"))
```


```{r}
longitudinal_dat %>%
  dplyr::group_by(school_id, year, var) %>%
  dplyr::summarise(n = dplyr::n(), .groups = "drop") %>%
  dplyr::filter(n > 1L)
```



```{r}
longitudinal_dat_wide <- longitudinal_dat %>% 
  pivot_wider(names_from = "var",
              values_from = "usage")
```

Wide longitudinal dataset created! Now just need to check the counts and values to make sure there was nothing wonky in the merging and everything is still accurate.

```{r}
# save(solo_tags, tags, longitudinal_dat_wide, longitudinal_dat, file = here::here("data", "longitudinal_tags.RData"))
# saving only wide and long as CSV file - J update 12.11.23
export(longitudinal_dat_wide, "data/longitudinal/tags-wide.csv")
export(longitudinal_dat, "data/longitudinal/tags-long.csv")
```

## Dataset verification
10/30
```{r}
longitudinal_dat %>% 
  count(year, var) %>% 
  group_by(year) %>% 
  summarise(n = n()) # looks like 2019 and 2022 both have one less than what is in the visual Janette and Merly made

#let's see what's missing
```
```{r eval=FALSE, include=FALSE}
# Ran the following code for each year: 2019, 2021, 2022, 2023

check_yr <- longitudinal_dat %>% 
  count(year, var) %>% 
  group_by(year) %>% 
  filter(year == "2023")

tags_check <- tags %>% 
  pivot_longer(cols = c(starts_with("equivalent_")),
               names_to = "year",
               values_to = "count") %>% 
  filter(count != 0) %>% 
  group_by(year) %>% 
  filter(year == "equivalent_variable_name_2023") %>% 
  select(current_variable_name, full_tag_name)

ccc <- left_join(tags_check, check_yr, by = join_by(current_variable_name == var))
```

In 2019, practices_portfolios_exhibitions is missing from longitudinal dataset.

In 2022, practices_service_learning is missing from longitudinal dataset.

Both 2021 and 2023 line up perfectly.

```{r}
use_long <- longitudinal_dat %>% 
  count(year, var) %>% 
  group_by(year)
```


```{r}
use_long %>% 
  ggplot(aes(x = year, y = n, group = var)) +
  geom_point(aes(color = var, alpha = 0.5)) +
  geom_line(aes(color = var, alpha = 0.5)) +
  theme_linedraw() +
  theme(legend.position = "none")
```


```{r message=FALSE, warning=FALSE, fig.width=20, fig.height=10}
all_use <- use_long %>% 
  ggplot(aes(x = year, y = n)) +
  geom_point(aes(color = var, alpha = 0.5)) +
  geom_line(aes(color = var, alpha = 0.5)) +
  theme_linedraw() +
  theme(legend.position = "none") +
  facet_wrap(~var) +
  theme(strip.text = element_text(size = 8, margin = margin()))

ggsave(filename = here("output", "all_tags_over_years.png"), width = 20, height = 10)
```

```{r}
library(plotly)

by_year <- use_long %>% 
  ggplot(aes(x = year, y = n, group = var)) +
  geom_point(aes(color = var, alpha = 0.5)) +
  geom_line(aes(color = var, alpha = 0.5)) +
  theme_linedraw() +
  theme(legend.position = "none")

ggplotly(by_year)
```


Tag correlations

```{r}
cor_tags <- cor(tags)
cor_tags <- cbind(tag = rownames(cor_tags), cor_tags)

write.csv(cor_tags, "correlations_2023.csv")
```

Separate all_use viz by cluster

## Educational justice and holistic student support

Social justice focus
Culturally responsive practices
Anti-racist practices
Restorative practices
Hiring for equity and inclusion values

## Postsecondary pathways and the world outside school

Career preparation and work-based learning
Students earn industry credentials
Assessments for career readiness
Community and business partnerships
Extended learning opportunities

## Deeper learning for mastery

Interdisciplinary 
Competency/mastery based education
Multi-age classrooms
Design thinking process
Makerspace

## Individualized and blended learning

Blended learning 
Interoperable data from multiple technologies
Students access their own data
Students move at their own pace
Student-led goal setting

```{r}
all_years <- use_long %>% 
  group_by(var) %>% 
  count(var) %>%
  filter(n == 4) %>% 
  select(-n)

all_years_viz <- inner_join(all_years, use_long)
```

```{r message=FALSE, warning=FALSE, fig.width=20, fig.height=10}

all_years_viz$year <- as.numeric(all_years_viz$year)
all_years_viz$var <- substring(all_years_viz$var, 11)

all_use_four_years <- all_years_viz %>% 
  ggplot(aes(x = year, y = n)) +
  geom_point(aes(color = var, alpha = 0.5)) +
  geom_line(aes(color = var, alpha = 0.5)) +
  theme_linedraw() +
  theme(legend.position = "none") +
  facet_wrap(~var) +
  theme(strip.text = element_text(size = 10, margin = margin()))

ggsave(filename = here("output", "tags_used_all_years.png"), width = 20, height = 10)
```


# Janette data checks 12.11.23

```{r}
# clear environment
rm(list = ls())
# read in data
dat <- import(here("data/longitudinal", "tags-wide.csv"))
# Data needs to be in 0/1/NA format - modify
dat <- dat %>% 
  mutate_all(~ ifelse(. == "", 0, .))
#create summary view for y/n in each year and N
long <- import(here("data/longitudinal", "tags-long.csv"))
# usage is not the right type - it's a character and needs to be numeric
long <- long %>% 
  mutate(usage = as.numeric(usage))
test <- long %>% 
  select(var, year, usage) %>% 
  group_by(year, var) %>% 
  summarize(n = sum(usage, na.rm = TRUE))
# totals look good
# years look good - note there's no 0s for years where tag wasn't used; tag just isn't there for long format
# portfolios exhibitions merge looks good
# resave wide/long after introducing 0/1 and turning usage into numeric
rio::export(dat, "data/longitudinal/tags-wide.csv")
rio::export(long, "data/longitudinal/tags-long.csv")
```

