---
title: "School participation across the year"
author: "Merly Klaas"
date: "2023-09-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(rio, here, dplyr,tidyverse,lubridate, ggplot2,viridis)
```

I did the initial cleaning in excel: Updating school name typo with school_id as the reference. There are many occurrences of school name inconsistencies with the same school id across the year. Now, each school_id corresponds to one identical school name. I then imported the cleaned dataset here. 

```{r}
school <- import(here("data", "long_school.csv"))

```

Pivot wider data to count for total school participation 
```{r}
school_wide <- school %>% 
  group_by(school_id,school_name, year) %>% 
  count(school_name) %>% 
  pivot_wider(names_from = year, values_from = n) %>% 
  mutate(total = rowSums(across(c("2019","2021", "2022", "2023")), na.rm=TRUE)) %>% 
  select(school_id, school_name, "2019","2021", "2022", "2023", total)
```


## Overall Number of school participated each year

*In 2023 we have the largest number of school participated (n=251), while 2022 is the least number of school participation (n=161).*


```{r}
school_wide %>% 
pivot_longer(cols=c("2019","2021", "2022", "2023", total),
               names_to = "year",
               values_to = "count") %>% 
    drop_na() %>% 
  group_by(year) %>% 
  count() %>% 
  ggplot()+
  geom_col(aes(year, n), fill= "#1A4C81") +
   geom_text(aes(year, n,label = n), vjust = -0.5)
```


## Schools' repeated participation trends across the year

However, in terms of schools regular participation, out of 503 schools ever participated in Canopy project, we only have 27 of schools  participated in all 4 year, 51 school participated in 3 time points, 131 schools in 2 time points, and 294 schools only 1 time point. 

```{r}

table(school_wide$total)
school_wide %>% 
 ggplot(aes(x=as.factor(total))) +
  geom_bar(fill="#1A4C81") + 
  geom_text(aes(y = ((..count..)), label = (..count..)), stat = "count", vjust = -0.5)

```

## Each year and the accumulated school participation 

Here we can see the trend of each school participation from  2019 - 2023 (gray = not participated, yellow = participated), and the far right is the total count of year participated. Color map confirmed the least participation happened in 2022. 

```{r}
school_wide %>% 
  mutate(n = total) %>% 
    pivot_longer(cols = c("2019","2021", "2022", "2023", total),
               names_to = "year",
               values_to = "count") %>%  
   ggplot() +
  geom_tile(aes(year, y=fct_reorder(school_name, n), fill=count))+
   scale_fill_viridis(discrete=FALSE, direction=-1)+
  labs(y= "Nominated Schools")+
  theme(axis.text.y = element_blank())
```

```{r}


```


