---
title: "initial-tags"
author: "Merly Klaas"
date: "2023-10-06"
output: 
  html_document:
    toc_float: yes
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
pacman::p_load(rio, here, dplyr,tidyverse,lubridate, ggplot2,viridis)
```

```{r}
tags <- 
  import(here("data", "Tags 2019-2023 OLD.csv")) %>% # NOTE: this dataset is old and current should be used
   mutate(across(starts_with("y_"),
                ~ case_when(is.na(.) ~ 0, TRUE ~ 1)),
         total = rowSums(across(starts_with("y_")), na.rm=TRUE))
```


## Number of tags used across the year

In total, there are 137 different tags used in the span of 4 times surveys. 87 tags were used in 2019, 91 tags were used in 2021, 72 tags were used in 2022, and 73 tags were used in 2023.

```{r}
tags %>% 
    pivot_longer(cols = c(starts_with("y_"), "total"),
               names_to = "year",
               values_to = "count")%>% 
  filter(count != 0) %>% 
  group_by(year) %>% 
  count() %>% 
  ggplot()+
  geom_col(aes(year, n), fill= "#1A4C81") +
   geom_text(aes(year, n,label = n), vjust = -0.5)
```

## Count of repeated tags used

42 tags were used in all 4 time point, 23 tags were used in 3 time points, 14 tags were used in 2 time points, and 58 tags were only used once. 

```{r}

tags %>% 
ggplot(aes(x=as.factor(total))) +
  geom_bar(fill="#1A4C81") + 
  geom_text(aes(y = ((..count..)), label = (..count..)), stat = "count", vjust = -0.5)+
  labs(x= "Times used",
       y= "Number of Variables")

```

## Visual of consistent year-over-year tags used in the Canopy survey

```{r fig.height=10, fig.width=7}
tags %>% 
   mutate(n = total) %>% 
    pivot_longer(cols = c(starts_with("y_"), total),
               names_to = "year",
               values_to = "count") %>%  
  mutate(count = case_when(count == 0 ~ NA, TRUE ~count),
         year = fct_relevel(year,"y_2019", "y_2021","y_2022","y_2023", "total" )) %>% 
   ggplot() +
  geom_tile(aes(year, y=fct_reorder(variable_name, n), fill=count))+
   scale_fill_viridis(discrete=FALSE, direction=-1)+
scale_x_discrete(expand = c(0, 0))+
  labs(y = "Tags Name") + 
  theme(axis.text=element_text(size=7))
```

## Updates to visuals post-fix
Janette 10.27.23

## Number of tags used across the year

In total, there are 130 different tags used in the span of 4 times surveys. 87 tags were used in 2019, 91 tags were used in 2021, 72 tags were used in 2022, and 73 tags were used in 2023.

```{r}
load(here("data/longitudinal_tags.RData"))
tags <- tags %>% 
  mutate(y_2023 = case_when(
    !is.na(equivalent_variable_name_2023) ~ 1
  ),
  y_2022 = case_when(
    !is.na(equivalent_variable_name_2022) ~ 1
  ),
  y_2021 = case_when(
    !is.na(equivalent_variable_name_2021) ~ 1
  ),
  y_2019 = case_when(
    !is.na(equivalent_variable_name_2019) ~ 1
  )) %>%
  select(current_variable_name, full_tag_name, starts_with("y_")) %>% 
  group_by(full_tag_name) %>% 
  mutate(total = sum(y_2023, y_2022, y_2021, y_2019, na.rm = TRUE)) 

tags %>% pivot_longer(cols = c(starts_with("y_"), "total"),
               names_to = "year",
               values_to = "count") %>% 
  filter(count != 0) %>% 
  group_by(year) %>% 
  count() %>% 
  ggplot()+
  geom_col(aes(year, n), fill= "#1A4C81") +
  geom_text(aes(year, n,label = n), vjust = -0.5)
```

## Count of repeated tags used

49 tags were used in all 4 time point, 17 tags were used in 3 time points, 12 tags were used in 2 time points, and 52 tags were only used once. 

```{r}
tags %>% 
ggplot(aes(x=as.factor(total))) +
  geom_bar(fill="#1A4C81") + 
  geom_text(aes(y = ((..count..)), label = (..count..)), stat = "count", vjust = -0.5)+
  labs(x= "Times used",
       y= "Number of Variables")
```

## Visual of consistent year-over-year tags used in the Canopy survey

```{r fig.height=10, fig.width=7}
tags %>% 
   mutate(n = total) %>% 
    pivot_longer(cols = c(starts_with("y_"), total),
               names_to = "year",
               values_to = "count") %>%  
  mutate(count = case_when(count == 0 ~ NA, TRUE ~count),
         year = fct_relevel(year,"y_2019", "y_2021","y_2022","y_2023", "total" )) %>% 
   ggplot() +
  geom_tile(aes(year, y=fct_reorder(full_tag_name, n), fill=count))+
   scale_fill_viridis(discrete=FALSE, direction=-1)+
scale_x_discrete(expand = c(0, 0))+
  labs(y = "Tags Name") + 
  theme(axis.text=element_text(size=7))
```
