---
title: "Catalysts for Innovation"
author: "Anwesha Guha"
date: "2024-03-28"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
pacman::p_load(tidyverse, here, rio, DT, ggcorrplot, psych, parameters)

#read in catalyst data
dat <- import(here("data/longitudinal", "longitudinal_data.csv"))

#read in tag data
tag_dat <- import(here("data/longitudinal", "tags-long.csv"))

#import tag labels
labels <- import(here("data/longitudinal", "tag-labels.csv"))
```


```{r}
load(here("data/2024 data", "complete_canopy_2024.RData"))
```

## Reflections

For `catalyst_key`, here are some notable relationships/correlations (0.25 and above).

_2021_

* cutting_edge: flexible_facilities
* inequities: culturally_responsive, hiring_equity, social_justice, anti-racist, design-marginalized

_2024_

* covid: enriched_virtual
* stakeholders: adult_wellness

For just `catalyst_`, here are some notable relationships/correlations (0.25 and above).

_2021_

* cutting_edge: flexible_facilities
* inequities: culturally_responsive, social_justice, hiring_equity,  anti-racist, design-marginalized

_2024_

* cutting_edge: interdisciplinary
* demographics: translanguaging
* inequities: hiring_equity

## Correlation Analysis

### Catalyst Correlations with 2021 data

```{r}
catalyst_dat_2021 <- dat %>% 
  filter(year == 2021) %>% 
  select(school_id, starts_with("catalyst_"))

tag_dat_2021 <- tag_dat %>% 
  filter(year == 2021) %>% 
  select(-year)

# Combine catalyst and tag files.
dat_2021 <- full_join(catalyst_dat_2021, tag_dat_2021)


# Calculate correlations for practice by catalyst. 
results_catalyst_2021 <- dat_2021 %>%
  select(starts_with("catalyst_"), usage, var, -catalyst_other_text, -starts_with("catalyst_key")) %>%
  group_by(var) %>%
  summarise(across(starts_with("catalyst_"), ~round(cor(.x, usage, use="complete.obs"), 2)))

results_catalyst_2021 %>% 
  datatable()
```

```{r}
results_catalyst_key_2021 <- dat_2021 %>%
  select(starts_with("catalyst_key"), usage, var) %>%
  group_by(var) %>%
  summarise(across(starts_with("catalyst_"), ~round(cor(.x, usage, use="complete.obs"), 2)))

results_catalyst_key_2021 %>% 
  datatable()
```


### Catalyst Correlations with 2024 data

```{r}
catalyst_dat_2024 <- dat %>% 
  filter(year == 2024) %>% 
  select(school_id, starts_with("catalyst_"))

tag_dat_2024 <- tag_dat %>% 
  filter(year == 2024) %>% 
  select(-year)

# Combine catalyst and tag files.
dat_2024 <- full_join(catalyst_dat_2024, tag_dat_2024)


# Calculate correlations for practice by catalyst. 
results_catalyst_2024 <- dat_2024 %>%
  select(starts_with("catalyst_"), usage, var, -catalyst_other_text, -starts_with("catalyst_key")) %>%
  group_by(var) %>%
  summarise(across(starts_with("catalyst_"), ~round(cor(.x, usage, use="complete.obs"), 2)))

results_catalyst_2024 %>% 
  datatable()
```

```{r}
results_catalyst_key_2024 <- dat_2024 %>%
  select(starts_with("catalyst_key"), usage, var) %>%
  group_by(var) %>%
  summarise(across(starts_with("catalyst_"), ~round(cor(.x, usage, use="complete.obs"), 2)))

results_catalyst_key_2024 %>% 
  datatable()
```

### Catalyst Update -- just 2024 data

Which catalysts are unique to 2024?
```{r}
catalyst_dat_only_2024 <- variables %>% 
  select(school_id, starts_with("catalyst_"))

setdiff(colnames(catalyst_dat_only_2024), colnames(catalyst_dat_2024))
```

* absence (chronic absenteeism)
* mental_health (mental health)

```{r}
tag_dat_2024 <- tag_dat %>% 
  filter(year == 2024) %>% 
  select(-year)

# Combine catalyst and tag files.
dat_only_2024 <- full_join(catalyst_dat_only_2024, tag_dat_2024)


# Calculate correlations for practice by catalyst. 
results_catalyst_only_2024 <- dat_only_2024 %>%
  select(starts_with("catalyst_"), usage, var, -catalyst_other_text, -starts_with("catalyst_key")) %>%
  group_by(var) %>%
  summarise(across(starts_with("catalyst_"), ~round(cor(.x, usage, use="complete.obs"), 2)))

results_catalyst_only_2024 %>% 
  datatable()
```


```{r}
results_catalyst_key_only_2024 <- dat_only_2024 %>%
  select(starts_with("catalyst_key"), usage, var) %>%
  group_by(var) %>%
  summarise(across(starts_with("catalyst_"), ~round(cor(.x, usage, use="complete.obs"), 2)))

results_catalyst_key_only_2024 %>% 
  datatable()
```

### Catalyst Correlations (with each other)

```{r}
catalyst_cor <- cor(catalyst_dat_2024[2:10], use = "pairwise.complete.obs")  # exclude other columns, like id, key, & other

corrplot::corrplot(catalyst_cor, method = "shade", type = "lower")
```

### Catalysts by Clusters

For 2024:
```{r}
old_clusters <- import(here("data/clusters_through_2024.csv"))
cluster_dat <- left_join(tag_dat_2024, old_clusters)

dat_only_2024 <- full_join(catalyst_dat_only_2024, cluster_dat)

results_catalyst_only_2024 <- dat_only_2024 %>%
  select(starts_with("catalyst_"), usage, cluster, -catalyst_other_text, -starts_with("catalyst_key")) %>%
  group_by(cluster) %>%
  summarise(across(starts_with("catalyst_"), ~round(cor(.x, usage, use="complete.obs"), 2))) %>% 
  pivot_longer(!cluster, names_to = "catalyst", values_to = "correlation") %>% 
  pivot_wider(names_from = cluster,
              values_from = correlation)

results_catalyst_only_2024 %>% 
  datatable()
```


## Changes over Time

```{r}
source(here("scripts/branding.R"))
```

Evolution of catalyst questions (changes to questions from 2021):

* Desire to be on the cutting edge and demonstrate what’s possible -- *same*
* External catalyst events (e.g., a partnership, a visit to another school, an unexpected source of funding) -- *same*
* Lack of student agency -- *same*
* Lack of teacher agency -- *same*
* Change in demographics in our community -- *"in our community" added*
* Negative factors inside the school (e.g., toxic culture, low achievement) -- *same*
* Stakeholder demand or advocacy (e.g., from the community, district leadership, teachers, or state agency) -- *same*
* Systemic inequities (e.g., racism, poverty) -- *same*
* Chronic absenteeism -- *ADDED*
* Learning disruptions due to COVID -- *changed from "School building closures related to COVID-19"*
* Mental health concerns -- *ADDED*
* Other (please specify) __________ -- *same*

```{r}
catalyst_dat_long <- dat %>% 
  filter(year == 2021 | year == 2024) %>% 
  select(school_id, year, starts_with("catalyst_")) %>% 
  select(-starts_with("catalyst_key"), -starts_with("catalyst_other"))

catalyst_key_dat_long <- dat %>% 
  filter(year == 2021 | year == 2024) %>% 
  select(school_id, year, starts_with("catalyst_key")) %>% 
  select(-starts_with("catalyst_key_other"))
```


### Across Schools

For just catalyst: 
```{r}
catalyst_dat <- catalyst_dat_long %>% 
  pivot_longer(cols = -c(school_id, year),
               names_to = "catalyst",
               values_to = "usage",
               names_pattern = "^catalyst_(.*)$")

catalyst_use <- catalyst_dat %>%
  group_by(catalyst, year) %>%
  summarize(total_use = sum(usage))
```

```{r}
library(ggrepel)
label_positions <- catalyst_use %>% 
  group_by(catalyst) %>% 
  summarize(year = 2021, total_use = first(total_use))

catalyst_use %>% 
  ggplot(aes(x = year, y = total_use, color = catalyst)) +
  geom_line() +
  geom_text_repel(data = label_positions, aes(label = catalyst)) +
  scale_x_continuous(breaks = c(2021, 2024)) +
  labs(x = "Year",
       y = "Number of Times Selected",
       title = "Catalyst Selection by Year") +
  theme(legend.position = "none")
```

*Note:* Demographics and teacher_agency are increasingly selected as a catalyst for innovation, though selection for other catalyst categories on the whole have decreased.

For catalyst_key:
```{r}
catalyst_key_dat <- catalyst_key_dat_long %>% 
  pivot_longer(cols = -c(school_id, year),
               names_to = "catalyst_key",
               values_to = "usage",
               names_pattern = "^catalyst_key_(.*)$")

catalyst_key_use <- catalyst_key_dat %>%
  group_by(catalyst_key, year) %>%
  summarize(total_use = sum(usage))
```

```{r}
label_positions <- catalyst_key_use %>% 
  group_by(catalyst_key) %>% 
  summarize(year = 2021, total_use = first(total_use))

catalyst_key_use %>% 
  ggplot(aes(x = year, y = total_use, color = catalyst_key)) +
  geom_line() +
  geom_text_repel(data = label_positions, aes(label = catalyst_key)) +
  scale_x_continuous(breaks = c(2021, 2024)) +
  labs(x = "Year",
       y = "Number of Times Selected",
       title = "Catalyst Key Selection by Year") +
  theme(legend.position = "none")
```


### Within Schools

For just catalyst:
```{r}
repeat_sch_catalyst <- catalyst_dat_long[duplicated(catalyst_dat_long$school_id) | duplicated(catalyst_dat_long$school_id, fromLast = TRUE),]

repeat_sch_catalyst_long <- repeat_sch_catalyst %>% 
  pivot_longer(cols = -c(school_id, year),
               names_to = "catalyst",
               values_to = "usage",
               names_pattern = "^catalyst_(.*)$")

repeat_sch_catalyst_long <- repeat_sch_catalyst_long %>% 
  pivot_wider(names_from = "year",
              values_from = "usage")

repeat_sch_catalyst_long <- repeat_sch_catalyst_long %>% 
  mutate(status = case_when(`2021` == 1 & `2024` == 1 ~ "selected both years",
                            `2021` == 1 & `2024` == 0 ~ "dropped",
                            `2021` == 0 & `2024` == 1 ~ "added",
                            `2021` == 0 & `2024` == 0 ~ "never selected"))
```

```{r}
catalyst_status_dat <- repeat_sch_catalyst_long %>% 
  group_by(catalyst, status) %>% 
  summarise(n = n()) %>% 
  pivot_wider(names_from = "status",
              values_from = "n")

catalyst_status_dat %>%  datatable()
```


```{r}
repeat_cat_viz <- repeat_sch_catalyst_long %>% 
  count(catalyst, status)

repeat_cat_viz %>% 
  ggplot(aes(x = catalyst, y = n, fill = status)) +
  geom_col(position = "dodge") +
  coord_flip() +
  scale_fill_manual(values = transcend_cols) +
  labs(title = "Catalyst Selection for Schools with Both Years of Data",
       x = "Catalyst",
       y = "Count",
       legend = "Status")
```

```{r}
repeat_cat_viz <- repeat_sch_catalyst_long %>% 
  count(catalyst, status) %>% 
  filter(status != "never selected")

repeat_cat_viz %>% 
  ggplot(aes(x = catalyst, y = n, fill = status)) +
  geom_col(position = "dodge") +
  coord_flip() +
  scale_fill_manual(values = transcend_cols) +
  labs(title = "Catalyst Selection for Schools with Both Years of Data",
       x = "Catalyst",
       y = "Count",
       legend = "Status")
```

Cutting edge was byfar the most selected catalyst for schools with two years of data.

#### Just one catalyst selected

```{r}
single_use_cat_dat <- catalyst_dat %>% 
  filter(usage == 1 & year == 2024) %>% 
  group_by(school_id) %>% 
  filter(n() == 1) %>% 
  ungroup()

single_use_cat_dat$catalyst <- factor(single_use_cat_dat$catalyst, levels = names(sort(table(single_use_cat_dat$catalyst), decreasing = TRUE))) #sort decreasing

single_use_cat_dat %>% 
  ggplot(aes(x = catalyst)) +
  geom_bar(aes(fill = catalyst)) +
  coord_flip() +
  scale_fill_manual(values = transcend_cols2) +
  labs(title = "Schools that Selected Only One Catalyst in 2024",
       x = "Catalyst",
       y = "Count") +
  theme(legend.position = "none")
```


```{r}
single_use_cat_dat2 <- catalyst_dat %>% 
  filter(usage == 1 & year == 2021) %>% 
  group_by(school_id) %>% 
  filter(n() == 1) %>% 
  ungroup()

single_use_cat_dat_all <- bind_rows(single_use_cat_dat, single_use_cat_dat2)

single_use_cat_dat_all$catalyst <- factor(single_use_cat_dat_all$catalyst, levels = names(sort(table(single_use_cat_dat_all$catalyst), decreasing = TRUE))) #sort decreasing

single_use_cat_dat_all %>% 
  ggplot(aes(x = catalyst)) +
  geom_bar(aes(fill = catalyst)) +
  coord_flip() +
#  scale_fill_manual(values = transcend_cols) +
  labs(title = "Schools that Selected Only One Catalyst across Years",
       x = "Catalyst",
       y = "Count") +
  theme(legend.position = "none")
```

Interestingly, basically no schools only selected one catalyst both years. There was only one school, school **24**, that was the anomaly, and they switched from *inequities* to *demographics*.

For catalyst_key:
```{r}
repeat_sch_catalyst_key <- catalyst_key_dat_long[duplicated(catalyst_key_dat_long$school_id) | duplicated(catalyst_key_dat_long$school_id, fromLast = TRUE),]

repeat_sch_catalyst_key_long <- repeat_sch_catalyst_key %>% 
  pivot_longer(cols = -c(school_id, year),
               names_to = "catalyst_key",
               values_to = "usage",
               names_pattern = "^catalyst_key_(.*)$")

repeat_sch_catalyst_key_long <- repeat_sch_catalyst_key_long %>% 
  pivot_wider(names_from = "year",
              values_from = "usage")

repeat_sch_catalyst_key_long <- repeat_sch_catalyst_key_long %>% 
  mutate(status = case_when(`2021` == 1 & `2024` == 1 ~ "selected both years",
                            `2021` == 1 & `2024` == 0 ~ "dropped",
                            `2021` == 0 & `2024` == 1 ~ "added",
                            `2021` == 0 & `2024` == 0 ~ "never selected"))
```

```{r}
catalyst_key_status_dat <- repeat_sch_catalyst_key_long %>% 
  group_by(catalyst_key, status) %>% 
  summarise(n = n()) %>% 
  pivot_wider(names_from = "status",
              values_from = "n")

catalyst_key_status_dat %>%  datatable()
```

```{r}
repeat_catkey_viz <- repeat_sch_catalyst_key_long %>% 
  count(catalyst_key, status)

repeat_catkey_viz %>% 
  ggplot(aes(x = catalyst_key, y = n, fill = status)) +
  geom_col(position = "dodge") +
  coord_flip() +
  scale_fill_manual(values = transcend_cols) +
  labs(title = "Catalyst Key Selection for Schools with Both Years of Data",
       x = "Catalyst",
       y = "Count",
       legend = "Status")
```

```{r}
repeat_catkey_viz <- repeat_sch_catalyst_key_long %>% 
  count(catalyst_key, status) %>% 
  filter(status != "never selected")

repeat_catkey_viz %>% 
  ggplot(aes(x = catalyst_key, y = n, fill = status)) +
  geom_col(position = "dodge") +
  coord_flip() +
  scale_fill_manual(values = transcend_cols) +
  labs(title = "Catalyst Key Selection for Schools with Both Years of Data",
       x = "Catalyst",
       y = "Count",
       legend = "Status")
```

## Catalyst Groups

```{r}
cat_combinations <- catalyst_dat_2024 %>%
  select(1:10) %>%   #just selecting catalyst vars
  rowwise() %>%
  mutate(combination = toString(names(.)[which(unlist(c_across(where(is.numeric))) == 1)])) %>%
  ungroup() %>% 
  count(combination, sort = TRUE) %>% 
  slice(-3) #missing values, just hard coding bc NA wasn't working for some reason
```

```{r fig.height=7, fig.width=15}
ggplot(cat_combinations[1:5, ], aes(x = reorder(combination, n), y = n, fill = combination)) +
  geom_col() +
  labs(title = "Top Catalyst Combinations in 2024", 
       x = "Combination", 
       y = "Frequency") +
  coord_flip() +
  scale_fill_manual(values = transcend_cols2) +
  theme(legend.position = "none",
        axis.text = element_text(size = 15))
```

Compare to 2021:

```{r}
cat_combinations <- catalyst_dat_2021 %>%
  select(1:10) %>%   #just selecting catalyst vars
  rowwise() %>%
  mutate(combination = toString(names(.)[which(unlist(c_across(where(is.numeric))) == 1)])) %>%
  ungroup() %>% 
  count(combination, sort = TRUE) %>% 
  slice(-7) #missing values, just hard coding bc NA wasn't working for some reason
```

```{r fig.height=7, fig.width=15}
ggplot(cat_combinations[1:5, ], aes(x = reorder(combination, n), y = n, fill = combination)) +
  geom_col() +
  labs(title = "Top Catalyst Combinations in 2021", 
       x = "Combination", 
       y = "Frequency") +
  coord_flip() +
  scale_fill_manual(values = transcend_cols2) +
  theme(legend.position = "none",
        axis.text = element_text(size = 15))
```


