---
title: "Cluster Trends"
author: "Janette Avelar"
date: '2024-03-07'
output: 
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warnings = FALSE, message = FALSE)
pacman::p_load(tidyverse, here, rio, DT, parameters, ggrepel, plotly, broom, lme4, gridExtra)
#read in data
dat <- import(here("data/longitudinal", "tags-long.csv"))
#read in tag labels
labels <- import(here("data", "tag_labels.csv"))
#read in cluster information
cluster <- import(here("data/longitudinal", "tag_clusters_longitudinal.csv")) %>% 
  janitor::clean_names()
```


```{r}
# start by creating cluster subsets
#subset values
deeper <- cluster %>% 
  filter(proposed_cluster_for_preliminary_24_analysis == "Deeper learning") %>% 
  select(label = tag) %>% 
  left_join(labels, by = "label") %>% 
  pull(var)
ed_justice <- cluster %>% 
  filter(proposed_cluster_for_preliminary_24_analysis == "Ed justice") %>% 
  select(label = tag) %>% 
  left_join(labels, by = "label") %>% 
  pull(var)
individualized <- cluster %>% 
  filter(proposed_cluster_for_preliminary_24_analysis == "Individualized") %>% 
  select(label = tag) %>% 
  left_join(labels, by = "label") %>% 
  pull(var)
postsecondary <- cluster %>% 
  filter(proposed_cluster_for_preliminary_24_analysis == "Postsecondary") %>% 
  select(label = tag) %>% 
  left_join(labels, by = "label") %>% 
  pull(var)
none <- cluster %>%
  filter(proposed_cluster_for_preliminary_24_analysis == "None" |
           proposed_cluster_for_preliminary_24_analysis == "None?") %>%
  select(label = tag) %>%
  left_join(labels, by = "label") %>%
  pull(var)
```

```{r}
### DATA FUNCTIONS ###
#set up repeating schools - 2+ years
repeaters_2 <- dat %>% 
  select(school_id, year) %>% 
  unique() %>% 
  mutate(rate = 1) %>% 
  pivot_wider(names_from = "year",
              values_from = "rate") %>% 
  rowwise() %>% 
  mutate(participate = sum(`2019`, `2021`, `2022`, `2023`, `2024`, na.rm = TRUE)) %>% 
  filter(participate > 1) %>% 
  pull(school_id)
#set up repeating schools - 3+ years
repeaters_3 <- dat %>% 
  select(school_id, year) %>% 
  unique() %>% 
  mutate(rate = 1) %>% 
  pivot_wider(names_from = "year",
              values_from = "rate") %>% 
  rowwise() %>% 
  mutate(participate = sum(`2019`, `2021`, `2022`, `2023`, `2024`, na.rm = TRUE)) %>% 
  filter(participate > 2) %>% 
  pull(school_id)
#set up repeating schools - all years
repeaters_all <- dat %>% 
  select(school_id, year) %>% 
  unique() %>% 
  mutate(rate = 1) %>% 
  pivot_wider(names_from = "year",
              values_from = "rate") %>% 
  rowwise() %>% 
  mutate(participate = sum(`2019`, `2021`, `2022`, `2023`, `2024`, na.rm = TRUE)) %>% 
  filter(participate == 5) %>% 
  pull(school_id)
#set up totals 
tots <- dat %>% 
  group_by(year) %>% 
  summarize(total = n_distinct(school_id))
tots_2 <- dat %>% 
  filter(school_id %in% repeaters_2) %>% 
  group_by(year) %>% 
  summarize(total = n_distinct(school_id))
tots_3 <- dat %>% 
  filter(school_id %in% repeaters_3) %>% 
  group_by(year) %>% 
  summarize(total = n_distinct(school_id))
tots_all <- dat %>% 
  filter(school_id %in% repeaters_all) %>% 
  group_by(year) %>% 
  summarize(total = n_distinct(school_id))
#set up 3+ year tags
rep_tags <- dat %>% 
  group_by(var) %>% 
  summarize(num_years = n_distinct(year)) %>% 
  ungroup() %>% 
  filter(num_years >= 3) %>% 
  select(var) %>% 
  unique() %>% 
  pull(var)
#prep data function across schools
prep_across <- function(data, cluster){
  dat <- data %>% 
    filter(var %in% cluster) %>% 
    group_by(var, year) %>% 
    summarize(n = sum(usage)) %>% 
    ungroup() %>% 
    left_join(tots, by = "year") %>% 
    mutate(pct = n/total) %>% 
    left_join(labels, by = "var") %>% 
    group_by(var) %>% 
    mutate(x_coord = max(year),
           y_coord = pct[which.max(year)])
  
  return(dat)
}
#prep data function within schools
prep_within <- function(data, cluster){
  dat <- data %>% 
    filter(var %in% cluster) %>% 
    filter(school_id %in% repeaters_2) %>% 
    arrange(school_id, year) %>% 
    group_by(school_id) %>%
    mutate(wave = dense_rank(year)) %>% 
    ungroup() %>% 
    arrange(school_id, year) %>%
    group_by(school_id, var) %>%
    mutate(change = usage - lag(usage)) %>% 
    ungroup() %>% 
    mutate(adds = ifelse(change == 1, 1, NA),
           drops = ifelse(change == -1, 1, NA)) %>% 
    group_by(var, year) %>% 
    summarize(`Overall rate` = sum(usage, na.rm = TRUE),
              `Average change` = sum(change, na.rm = TRUE),
              Adds = sum(adds, na.rm = TRUE),
              Drops = sum(drops, na.rm = TRUE)) %>% 
    ungroup() %>% 
    pivot_longer(cols = c(`Overall rate`, `Average change`, Adds, Drops),
                 names_to = "type",
                 values_to = "n") %>% 
    left_join(labels, by = "var") %>% 
    mutate(type = factor(type, levels = c("Overall rate", "Average change", "Adds", "Drops"))) %>% 
    #3.27.24 modification = remove 2019 & only present adds/drops
    filter(year > 2019, type == "Adds" | type == "Drops")
    return(dat)
}
#prep data function overall pct
prep_overall <- function(data, cluster, group, total){
  dat <- data %>% 
    filter(var %in% cluster) %>% 
    filter(school_id %in% group) %>% 
    group_by(var, year) %>% 
    summarize(n = sum(usage, na.rm = TRUE)) %>% 
    ungroup() %>% 
    left_join(total, by = "year") %>% 
    mutate(pct = n/total) %>% 
    left_join(labels, by = "var") %>% 
    group_by(var) %>% 
    mutate(x_coord = max(year),
           y_coord = pct[which.max(year)])

    return(dat)
}
prep_net <- function(data){
  data <- data %>% 
    pivot_wider(names_from = type,
                values_from = n) %>%
    group_by(var) %>% 
    mutate(net = Adds - Drops) %>% 
    ungroup() %>% 
    left_join(tots, by = "year") %>% 
    group_by(year, var, label) %>% 
    summarize(norm_net = net/total) %>% 
    ungroup() %>% 
    group_by(var) %>% 
    mutate(cum_norm_net = mean(norm_net)) %>% #average net change over time
    ungroup() %>% 
    mutate(type = case_when(
      cum_norm_net == max(cum_norm_net) ~ "Largest net change over time",
      cum_norm_net == min(cum_norm_net) ~ "Smallest net change over time",
      TRUE ~ "Average net change"
    ),
    type = factor(type, levels = c("Largest net change over time", "Smallest net change over time", "Average net change"))) #specify order for palette
}
prep_net_3 <- function(data){
  data <- data %>% 
    filter(var %in% rep_tags) %>% 
    mutate(max_cum = max(cum_norm_net),
           min_cum = min(cum_norm_net)) %>% 
  group_by(var) %>% 
  mutate(type = case_when(
         max_cum == cum_norm_net ~ "Largest net change over time",
         min_cum == cum_norm_net ~ "Smallest net change over time",
         TRUE ~ "Average net change"),
         type = factor(type, levels = c("Largest net change over time", "Smallest net change over time", "Average net change")))
}
```

```{r}
### PLOT FUNCTIONS ###
#plot function for across schools
cluster_plot_across <- function(data, x, y, labels) {
  #palette <- c("#BC2582", "#FFA630", "#FFDE42", "#99C24D", "#218380", "#D3B7D7")
  data <- data %>% 
    mutate(wrapped_label = str_wrap(get(labels), 40))
  
  labels <- data %>%
    select(label = wrapped_label, x_coord, y_coord) %>%
    unique()

  plot <- ggplot(data, aes_string(x = x, y = y, color = "label")) +
    geom_point() +
    geom_line() +
    theme_bw() +
    scale_x_continuous(limits = c(2019, 2027),
                       expand = c(0.01, 0.5),
                       breaks = unique(data$year)) + #expand 0.025, .3
    scale_y_continuous(limits = c(0, 1),
                       expand = c(0, 0),
                       labels = scales::percent_format(accuracy = 1)) +
    theme(legend.position = "none") +
    geom_text_repel(data = labels,
                    aes(label = label, x = x_coord, y = y_coord),
                    nudge_x = 0.5,
                    direction = "y",
                    size = 3,
                    box.padding = 0.3,
                    point.padding = 2
                    )
  
  return(plot)
}

cluster_plot_within <- function(data, facet) { #data, x, y, facet, group
  plot <- ggplot(data, aes_string(x = "year", y = "n")) + #aes_string(x = x, y = y, text = group)
    geom_line(aes(group = label), color = "gray", alpha = 0.5) + #group = group
    geom_point(aes(group = label), color = "gray", alpha = 0.5, size = 1) + #group = group
    geom_smooth(se = FALSE, method = "lm", color = "cornflowerblue", aes(group = 1)) + 
    #straight line before
    #geom_smooth(method = "lm", se = FALSE, color = "blue", aes(group = 1)) +
    theme_bw() +
    scale_x_continuous(name = "Year", expand = expansion(mult = c(.025, .3))) +
    theme(panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank(), legend.position = "none") +
    facet_wrap(as.formula(paste0("~", facet))) #+  #, scales = "free_y"
    #labs(x = x, y = y)

  return(plot)
}
net_change_plot <- function(data){
  palette <- c("#6cac8c", "#ff6f69","#d3d3d3")
  
  plot <- ggplot(data, aes(year, norm_net, color = type, group = var, text = label)) +
    geom_line(alpha = 0.8) +
    geom_point(alpha = 0.5, size = 1) +
    scale_color_manual(values = palette) +
    theme_bw() +
    theme(panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank(), legend.position = "none") +
    scale_x_continuous(expand = expansion(mult = c(0, 0))) +
    scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
    labs(x = "", y = "Normalized net change",
         subtitle = "Values above the dashed line indicate growth. Values below the dashed line indicate decline.") +
    geom_hline(aes(yintercept = 0), size = .25, linetype = "dashed")
  
  interactive <- ggplotly(plot, tooltip = c("year", "norm_net", "label"))
  
  return(interactive)
}
```

```{r}
#try plotly
#across plot
fancy_across <- function(data){
  plot <- plot_ly(data, x = ~year, y = ~pct, type = 'scatter', mode = 'lines+markers', text = ~label, hoverinfo = 'text+y') %>% 
    layout(yaxis = list(tickformat = ".0%"))
  return(plot)
}
```

# Deeper Learning Cluster

Tags:  
- assessments for deeper learning           
- competency/mastery-based education          
- competency framework            
- design thinking process          
- interdisciplinary            
- makerspace                  
- multiple opportunities to demonstrate mastery  
- project-based learning                  
- performance based assessment            
- place-based learning                      
- student-led conferences                
- students develop projects  

Guiding question: How have the rates of deeper learning tags changed each year since 2019?

### Across schools

```{r}
deep <- prep_across(dat, deeper)
#cluster_plot_across(deep, "year", "pct", "label")
fancy_across(deep)
```

### Within schools

Overall rate among repeat schools (2+ years):

```{r}
deep <- prep_overall(dat, deeper, repeaters_2, tots_2)
#cluster_plot_across(deep, "year", "pct", "label")
fancy_across(deep)
```

Overall rate among repeat schools (3+ years):

```{r}
deep <- prep_overall(dat, deeper, repeaters_3, tots_3)
fancy_across(deep)
```

Overall rate among repeat schools (5 years):

```{r}
deep <- prep_overall(dat, deeper, repeaters_all, tots_all)
fancy_across(deep)
```


```{r}
deep <- prep_within(dat, deeper)
plot <- cluster_plot_within(deep, "type") #deep, deep$year, deep$n, "type", deep$label
suppressWarnings({ggplotly(plot)})
```

### Net change

In the following graph I used net change (Adds - Drops) to calculate the rate of change for each tag. I then created a cumulative rate for each tag (average across years) to identify the highest and lowest rates of change in the cluster. The highest rate of change is colored in green and the lowest rate of change in red.

Note these plots are interactive--if you're curious about a particular line you can hover over one of it's points to see which tag it corresponds to. 

```{r}
deep <- prep_net(deep)
net_change_plot(deep)
```

### Net change 3+ year tags

```{r}
deep <- prep_net_3(deep)
net_change_plot(deep)
```


# Educational Justice Cluster

Tags:  
- adult wellness/SEL                    
- anti-racist practices                 
- assessments for social-emotional skills       
- teachers as co-leaders           
- co-leadership                       
- family and community support services         
- culturally responsive practices           
- design to meet needs of students who have been marginalized  
- hiring for equity and inclusion values                   
- all courses designed for inclusion               
- mental health services                       
- physical well being services          
- reallocation of resources for those most in need  
- restorative practices                    
- SEL curriculum                             
- SEL integration school-wide               
- social justice focus                  
- translanguaging                       
- trauma-informed practices  

### Across schools

```{r}
justice_dat <- prep_across(dat, ed_justice)
fancy_across(justice_dat)
```

### Within schools

Overall rate among repeat schools (2+ years):

```{r}
justice_dat <- prep_overall(dat, ed_justice, repeaters_2, tots_2)
fancy_across(justice_dat)
```

Overall rate among repeat schools (3+ years):

```{r}
justice_dat <- prep_overall(dat, ed_justice, repeaters_3, tots_3)
fancy_across(justice_dat)
```

Overall rate among repeat schools (all years):

```{r}
justice_dat <- prep_overall(dat, ed_justice, repeaters_all, tots_all)
fancy_across(justice_dat)
```


```{r}
justice_dat <- prep_within(dat, ed_justice)
plot <- cluster_plot_within(justice_dat, "type")
suppressWarnings({ggplotly(plot)})
```

### Net change

```{r}
justice_dat <- prep_net(justice_dat)
net_change_plot(justice_dat)
```

### Net change 3+ year tags

```{r}
justice_dat <- prep_net_3(justice_dat)
net_change_plot(justice_dat)
```

# Individualized Cluster

Tags:  
- accommodations provided to all students  
- blended learning              
- flex model  
- flipped classroom           
- interoperable data from multiple technologies  
- 1:1 mentoring  
- station rotation  
- students access their own data      
- self-paced learning  

### Across schools

```{r}
ind_dat <- prep_across(dat, individualized)
fancy_across(ind_dat)
```


### Within schools

Overall rate among repeat schools (2+ years):

```{r}
ind_dat <- prep_overall(dat, individualized, repeaters_2, tots_2)
fancy_across(ind_dat)
```

Overall rate among repeat schools (3+ years):

```{r}
ind_dat <- prep_overall(dat, individualized, repeaters_3, tots_3)
fancy_across(ind_dat)
```

Overall rate among repeat schools (all years):

```{r}
ind_dat <- prep_overall(dat, individualized, repeaters_3, tots_3)
fancy_across(ind_dat)
```


```{r}
ind_dat <- prep_within(dat, individualized)
plot <- cluster_plot_within(ind_dat, "type")
suppressWarnings({ggplotly(plot)})
```

### Net change

```{r}
ind_dat <- prep_net(ind_dat)
net_change_plot(ind_dat)
```

### Net change 3+ year tags

```{r}
ind_dat <- prep_net_3(ind_dat)
net_change_plot(ind_dat)
```

# Postsecondary Cluster

Tags:  
- à la carte model  
- assessments for career readiness  
- career prep and work-based learning  
- community-based organizations as co-leaders  
- industry-based partners as co-leaders  
- community and business partnerships  
- early college high school  
- extended learning opportunities     
- students earn industry credentials  

### Across schools

```{r}
post_dat <- prep_across(dat, postsecondary)
fancy_across(post_dat)
```

### Within schools

Overall rate among repeat schools (2+ years):

```{r}
post_dat <- prep_overall(dat, postsecondary, repeaters_2, tots_2)
fancy_across(post_dat)
```

Overall rate among repeat schools (3+ years):

```{r}
post_dat <- prep_overall(dat, postsecondary, repeaters_3, tots_3)
fancy_across(post_dat)
```

Overall rate among repeat schools (all years):

```{r}
post_dat <- prep_overall(dat, postsecondary, repeaters_all, tots_all)
fancy_across(post_dat)
```


```{r}
post_dat <- prep_within(dat, postsecondary)
plot <- cluster_plot_within(post_dat, "type")
suppressWarnings({ggplotly(plot)})
```

### Net change

```{r}
post_dat <- prep_net(post_dat)
net_change_plot(post_dat)
```

### Net change 3+ year tags

```{r}
post_dat <- prep_net_3(post_dat)
net_change_plot(post_dat)
```

### High Schools only

```{r}
# set up data prep
# ID high schools
hs <- import(here("data/longitudinal", "longitudinal_data.csv")) %>% 
  select(school_id, grades_high) %>% 
  filter(grades_high == 1) %>% 
  pull(school_id)
#set up repeating schools
hs_repeaters <- dat %>% 
  filter(school_id %in% hs) %>% 
  select(school_id, year) %>% 
  unique() %>% 
  mutate(rate = 1) %>% 
  pivot_wider(names_from = "year",
              values_from = "rate") %>% 
  rowwise() %>% 
  mutate(participate = sum(`2019`, `2021`, `2022`, `2023`, `2024`, na.rm = TRUE)) %>% 
  filter(participate > 1) %>% 
  pull(school_id)
#set up totals 
hs_tots <- dat %>% 
  filter(school_id %in% hs) %>% 
  group_by(year) %>% 
  summarize(total = n_distinct(school_id)) %>% 
  unique()
#prep data across schools
hs_across <- dat %>% 
  filter(school_id %in% hs) %>% 
    filter(var %in% postsecondary) %>% 
    group_by(var, year) %>% 
    summarize(n = sum(usage)) %>% 
    ungroup() %>% 
    left_join(hs_tots, by = "year", relationship = "many-to-many") %>% 
    mutate(pct = n/total) %>% 
    left_join(labels, by = "var") %>% 
    group_by(var) %>% 
    mutate(x_coord = max(year),
           y_coord = pct[which.max(year)])
#prep data within schools
hs_within <- dat %>% 
  filter(school_id %in% hs) %>% 
    filter(var %in% postsecondary) %>% 
    filter(school_id %in% hs_repeaters) %>% 
    arrange(school_id, year) %>% 
    group_by(school_id) %>%
    mutate(wave = dense_rank(year)) %>% 
    ungroup() %>% 
    arrange(school_id, year) %>%
    group_by(school_id, var) %>%
    mutate(change = usage - lag(usage)) %>% 
    ungroup() %>% 
    mutate(adds = ifelse(change == 1, 1, NA),
           drops = ifelse(change == -1, 1, NA)) %>% 
    group_by(var, year) %>% 
    summarize(`Overall rate` = sum(usage, na.rm = TRUE),
              `Average change` = sum(change, na.rm = TRUE),
              Adds = sum(adds, na.rm = TRUE),
              Drops = sum(drops, na.rm = TRUE)) %>% 
    ungroup() %>% 
    pivot_longer(cols = c(`Overall rate`, `Average change`, Adds, Drops),
                 names_to = "type",
                 values_to = "n") %>% 
    left_join(labels, by = "var") %>% 
    mutate(type = factor(type, levels = c("Overall rate", "Average change", "Adds", "Drops"))) %>% 
    #3.27.24 modification = remove 2019 & only present adds/drops
    filter(year > 2019, type == "Adds" | type == "Drops")
#prep data overall pct
hs_overall <- dat %>% 
    filter(var %in% postsecondary) %>% 
    filter(school_id %in% hs_repeaters) %>% 
    group_by(var, year) %>% 
    summarize(n = sum(usage, na.rm = TRUE)) %>% 
    ungroup() %>% 
    left_join(hs_tots, by = "year") %>% 
    mutate(pct = n/total) %>% 
    left_join(labels, by = "var") %>% 
    group_by(var) %>% 
    mutate(x_coord = max(year),
           y_coord = pct[which.max(year)])
```

### Across high schools

```{r}
fancy_across(hs_across)
```

### Within high schools

Overall rate among repeat high schools (2+ years):

```{r}
fancy_across(hs_overall)
```

```{r}
plot <- cluster_plot_within(hs_within, "type")
suppressWarnings({ggplotly(plot)})
```

### Net change

```{r}
hs_within <- prep_net(hs_within)
net_change_plot(hs_within)
```

### Net change 3+ year tags

```{r}
hs_within <- prep_net_3(hs_within)
net_change_plot(hs_within)
```

# None tags

### Across schools

```{r}
none_dat <- prep_across(dat, none)
fancy_across(none_dat)
```

### Within schools

Overall rate among repeat schools (2+ years):

```{r}
none_dat <- prep_overall(dat, none, repeaters_2, tots_2)
fancy_across(none_dat)
```

Overall rate among repeat schools (3+ years):

```{r}
none_dat <- prep_overall(dat, none, repeaters_3, tots_3)
fancy_across(none_dat)
```

Overall rate among repeat schools (all years):

```{r}
none_dat <- prep_overall(dat, none, repeaters_all, tots_all)
fancy_across(none_dat)
```


```{r}
none_dat <- prep_within(dat, none)
plot <- cluster_plot_within(none_dat, "type")
suppressWarnings({ggplotly(plot)})
```

### Net change

```{r}
none_dat <- prep_net(none_dat)
net_change_plot(none_dat)
```

### Net change 3+ year tags

```{r}
none_dat <- prep_net_3(none_dat)
net_change_plot(none_dat)
```

# Blended Learning Cluster

Reminder: This cluster is separate from the analysis - this was a starting point.

Tags:
- blended learning  
- à la carte model  
- flipped classroom  
- flex model  
- enriched virtual model  
- station rotation  

### Across schools

```{r}
# subset tags
blended_tags <- c("practices_blended_learning", "practices_a_la_carte", "practices_flipped_classroom", "practices_flex", "practices_enriched_virtual", "practices_station_rotation")
blend_dat <- prep_across(dat, blended_tags)
fancy_across(blend_dat)
```

### Within schools

Overall rate among repeat schools (2+ years):

```{r}
blend_dat <- prep_overall(dat, blended_tags, repeaters_2, tots_2)
fancy_across(blend_dat)
```

Overall rate among repeat schools (3+ years):

```{r}
blend_dat <- prep_overall(dat, blended_tags, repeaters_3, tots_3)
fancy_across(blend_dat)
```

Overall rate among repeat schools (all years):

```{r}
blend_dat <- prep_overall(dat, blended_tags, repeaters_all, tots_all)
fancy_across(blend_dat)
```

```{r}
blend_dat <- prep_within(dat, blended_tags)
plot <- cluster_plot_within(blend_dat, "type")
suppressWarnings({ggplotly(plot)})
```

### Net change

```{r}
blend_dat <- prep_net(blend_dat)
net_change_plot(blend_dat)
```

### Net change 3+ year tags

```{r}
blend_dat <- prep_net_3(blend_dat)
net_change_plot(blend_dat)
```

# All Clusters

### Net change

```{r}
merge_deep <- deep %>% 
  select(year, var, label, norm_net) %>% 
  mutate(cluster = "Deeper learning") %>% 
  group_by(year, cluster) %>% 
  summarize(cum_norm_net = mean(norm_net)) %>% 
  ungroup()
merge_ed_justice <- justice_dat %>% 
  select(year, var, label, norm_net) %>% 
  mutate(cluster = "Educational Justice") %>% 
  group_by(year, cluster) %>% 
  summarize(cum_norm_net = mean(norm_net)) %>% 
  ungroup()
merge_individualized <- ind_dat %>% 
  select(year, var, label, norm_net) %>% 
  mutate(cluster = "Individualized learning") %>% 
  group_by(year, cluster) %>% 
  summarize(cum_norm_net = mean(norm_net)) %>% 
  ungroup()
merge_postsecondary <- post_dat %>% 
  select(year, var, label, norm_net) %>% 
  mutate(cluster = "Postsecondary pathways") %>% 
  group_by(year, cluster) %>% 
  summarize(cum_norm_net = mean(norm_net)) %>% 
  ungroup()
merge_none <- none_dat %>% 
  select(year, var, label, norm_net) %>% 
  mutate(cluster = "No cluster") %>% 
  group_by(year, cluster) %>% 
  summarize(cum_norm_net = mean(norm_net)) %>% 
  ungroup()
merge_blended <- blend_dat %>% 
  select(year, var, label, norm_net) %>% 
  mutate(cluster = "Blended learning (alternate)") %>% 
  group_by(year, cluster) %>% 
  summarize(cum_norm_net = mean(norm_net)) %>% 
  ungroup()
all_clust <- bind_rows(merge_deep, merge_ed_justice, merge_individualized, merge_postsecondary, merge_none, merge_blended) %>% 
  rename("norm_net" = cum_norm_net) %>% 
  group_by(cluster) %>% 
  mutate(cum_norm_net = mean(norm_net)) %>% 
  ungroup() %>% 
  mutate(type = case_when(
         cum_norm_net == max(cum_norm_net) ~ "Largest net change over time",
         cum_norm_net == min(cum_norm_net) ~ "Smallest net change over time",
         TRUE ~ "Average net change"
        ),
        type = factor(type, levels = c("Largest net change over time", "Smallest net change over time", "Average net change")),
        label = cluster) %>% 
  rename("var" = cluster)
net_change_plot(all_clust)
rm(merge_blended, merge_deep, merge_individualized, merge_postsecondary, merge_ed_justice, merge_none)
```

### Net change (excluding alternate blended learning cluster)

```{r}
all_clust <- all_clust %>% 
  filter(var != "Blended learning (alternate)") %>% 
  mutate(max_cum = max(cum_norm_net),
           min_cum = min(cum_norm_net)) %>% 
  group_by(var) %>% 
  mutate(type = case_when(
         max_cum == cum_norm_net ~ "Largest net change over time",
         min_cum == cum_norm_net ~ "Smallest net change over time",
         TRUE ~ "Average net change"),
         type = factor(type, levels = c("Largest net change over time", "Smallest net change over time", "Average net change")))
net_change_plot(all_clust)
```


# First version graphs

### Across schools
Graph #1 looks at selection of tags across our whole sample, i.e., the number reflects the percentage of our total sample of schools that selected each tag. (AKA across schools)

One of the questions we were hoping to investigate with this graph is whether the selection of blended learning tags may be due to changes in our survey design through the introduction of parent/child tags. We do see a sharp decrease in child tags in 2023. However, the selection for the parent tag `blended learning` remains high. All school leaders that selected `blended learning` would have received a follow-up question about the type of blended learning they use with our child tags listed below. This makes me think that schools are still using blended learning, but perhaps did not see their approach to blended learning reflected in our child tags. (So what are they using...?)

```{r}
blended <- dat %>% 
  filter(var %in% blended_tags) %>%
  group_by(var, year) %>% 
  summarize(n = sum(usage)) %>% 
  ungroup()
tots <- dat %>% 
  group_by(year) %>% 
  summarize(total = n_distinct(school_id)) %>% 
  unique()
blended <- blended %>% 
  left_join(tots, by = "year") %>% 
  mutate(pct = n/total) %>% 
  left_join(labels, by = "var")
#build plot
#palette colors
palette <- c("#BC2582", "#FFA630","#FFDE42","#99C24D","#218380","#D3B7D7")
#labels
labs <- blended %>% 
  filter(year == 2024) %>% 
  select(label, x = year, y = pct)
#plot
blended %>% 
  ggplot(aes(year, pct, color = label)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  scale_color_manual(values = palette) +
  scale_x_continuous(expand = expansion(mult = c(.025, .3))) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  theme(legend.position = "none") +
  geom_text(data = labs, 
            aes(label = label, 
                x = x, 
                y = y),
            hjust = -.05)
```

### Across schools (no COVID)

What if we omit 2021 to account for the big jump?

```{r}
blended <- dat %>% 
  filter(var %in% blended_tags) %>%
  filter(year != 2021) %>% 
  group_by(var, year) %>% 
  summarize(n = sum(usage)) %>% 
  ungroup() %>% 
  left_join(tots, by = "year") %>% 
  mutate(pct = n/total) %>% 
  left_join(labels, by = "var")

#plot
blended %>% 
  ggplot(aes(year, pct, color = label)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  scale_color_manual(values = palette) +
  scale_x_continuous(expand = expansion(mult = c(.025, .3))) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  theme(legend.position = "none") +
  geom_text(data = labs, 
            aes(label = label, 
                x = x, 
                y = y),
            hjust = -.05)
```

### Within schools
Graph #2 looks at the rate of change for blended learning tags only among schools that participated for 2 or more years. (AKA within schools)

```{r}
#key for participating schools 2+ times
repeaters <- dat %>% 
  select(school_id, year) %>% 
  unique() %>% 
  mutate(rate = 1) %>% 
  pivot_wider(names_from = "year",
              values_from = "rate") %>% 
  rowwise() %>% 
  mutate(participate = sum(`2019`, `2021`, `2022`, `2023`, na.rm = TRUE)) %>% 
  filter(participate > 1) %>% 
  pull(school_id)
#data
blended <- dat %>% 
  #filter only blended tags
  filter(var %in% blended_tags) %>% 
  #filter only repeat schools
  filter(school_id %in% repeaters) %>% 
  #create wave variable
  arrange(school_id, year) %>% 
  group_by(school_id) %>%
  mutate(wave = dense_rank(year)) %>% 
  ungroup() %>% 
  #create change column
  arrange(school_id, year) %>%
  group_by(school_id, var) %>%
  mutate(change = usage - lag(usage)) %>% 
  ungroup() %>% 
  #create add/drop columns
  mutate(adds = ifelse(change == 1, 1, NA),
         drops = ifelse(change == -1, 1, NA)) %>% 
  #create totals
  group_by(var, year) %>% 
  summarize(`Overall rate` = sum(usage, na.rm = TRUE),
            `Average change` = sum(change, na.rm = TRUE),
            Adds = sum(adds, na.rm = TRUE),
            Drops = sum(drops, na.rm = TRUE)) %>% 
  ungroup() %>% 
  #pivot for plotting
  pivot_longer(cols = c(`Overall rate`, `Average change`, Adds, Drops),
               names_to = "type",
               values_to = "n") %>% 
  #labels for plotting
  left_join(labels, by = "var") %>% 
  #order for plotting
  mutate(type = factor(type, levels = c("Overall rate", "Average change", "Adds", "Drops")))

#palette for plot
palette <- c("gray80", "#f1c1af", "#6bd497", "#b94c4c")
#plot
blended %>% 
  ggplot(aes(year, n, color = type)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  scale_color_manual(values = palette) +
  scale_x_continuous(expand = expansion(mult = c(.025, .3))) +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()) +
  theme(legend.position = "bottom") +
  facet_wrap(~label) +
  labs(x = "Year",
       y = "Tag selection (N)",
       color = "Rate of change")
```

### Within schools (no COVID)
Again, what if we omit 2021?

```{r}
#data
blended <- dat %>% 
  #filter only blended tags
  filter(var %in% blended_tags) %>% 
  #filter only repeat schools
  filter(school_id %in% repeaters) %>% 
  #drop 2021 year
  filter(year != 2021) %>% 
  #create wave variable
  arrange(school_id, year) %>% 
  group_by(school_id) %>%
  mutate(wave = dense_rank(year)) %>% 
  ungroup() %>% 
  #create change column
  arrange(school_id, year) %>%
  group_by(school_id, var) %>%
  mutate(change = usage - lag(usage)) %>% 
  ungroup() %>% 
  #create add/drop columns
  mutate(adds = ifelse(change == 1, 1, NA),
         drops = ifelse(change == -1, 1, NA)) %>% 
  #create totals
  group_by(var, year) %>% 
  summarize(`Overall rate` = sum(usage, na.rm = TRUE),
            `Average change` = sum(change, na.rm = TRUE),
            Adds = sum(adds, na.rm = TRUE),
            Drops = sum(drops, na.rm = TRUE)) %>% 
  ungroup() %>% 
  #pivot for plotting
  pivot_longer(cols = c(`Overall rate`, `Average change`, Adds, Drops),
               names_to = "type",
               values_to = "n") %>% 
  #labels for plotting
  left_join(labels, by = "var") %>% 
  #order for plotting
  mutate(type = factor(type, levels = c("Overall rate", "Average change", "Adds", "Drops")))

#plot
blended %>% 
  ggplot(aes(year, n, color = type)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  scale_color_manual(values = palette) +
  scale_x_continuous(expand = expansion(mult = c(.025, .3))) +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()) +
  theme(legend.position = "bottom") +
  facet_wrap(~label) +
  labs(x = "Year",
       y = "Tag selection (N)",
       color = "Rate of change")
```

### 2023 Drop Investigation

Investigating drop in blended learning tags in 2023.

*From CW*: Could you do a quick tally of: 
a) how many schools in 2022 chose any of the child tags WITHOUT choosing "blended learning"

There were 21 instances where a school leader chose child tags but did not choose blended learning.

```{r}
dat %>% 
  filter(year == 2022) %>% 
  filter(var %in% blended_tags) %>% 
  pivot_wider(names_from = "var",
              values_from = "usage") %>% 
  group_by(school_id) %>% 
  mutate(children = case_when(
    practices_a_la_carte == 1 ~ 1,
    practices_flex == 1 ~ 1,
    practices_flipped_classroom == 1 ~ 1,
    practices_enriched_virtual == 1 ~ 1,
    practices_station_rotation == 1 ~ 1,
    TRUE ~ 0
  ),
  childless = case_when(
    practices_blended_learning == 1 & children == 0 ~ 1,
    TRUE ~ 0
  ),
  parentless = case_when(
    practices_blended_learning == 0 & children == 1 ~ 1,
    TRUE ~ 0
  )) %>% 
  ungroup() %>% 
  summarize(selected_children = sum(parentless),
            selected_parent_only = sum(childless)) %>% 
  datatable()
```


b) how many schools in 2023 chose "blended learning" as a parent tag but didn't choose any child tags?

There were 53 instances where a school leader chose blended learning but did not choose one of the child tags. (And 94 instances where they did choose child tags, for comparison.)

```{r}
dat %>% 
  filter(year == 2023) %>% 
  filter(var %in% blended_tags) %>% 
  pivot_wider(names_from = "var",
              values_from = "usage") %>% 
  group_by(school_id) %>% 
  mutate(children = case_when(
    practices_a_la_carte == 1 ~ 1,
    practices_flex == 1 ~ 1,
    practices_flipped_classroom == 1 ~ 1,
    practices_enriched_virtual == 1 ~ 1,
    practices_station_rotation == 1 ~ 1,
    TRUE ~ 0
  ),
  childless = case_when(
    practices_blended_learning == 1 & children == 0 ~ 1,
    TRUE ~ 0
  )) %>% 
  ungroup() %>% 
  summarize(selected_parent = sum(childless),
            selected_children = sum(children)) %>% 
  datatable()
```

# Models

```{r}
#clean up environment
rm(cluster, deep, justice_dat, tots, tots_2, tots_3, tots_all)
#create new cluster DV
mod_dat <- dat %>% 
  #add cluster groups
  mutate(cluster = case_when(
    var %in% deeper ~ "deeper",
    var %in% ed_justice ~ "ed_justice",
    var %in% individualized ~ "individualized",
    var %in% postsecondary ~ "postsecondary",
    var %in% none ~ "misc"
  )) %>% 
  #note: clusters only outlined for 2024 tags... do we want to modify?
  #for now, drop missing cluster
  filter(!is.na(cluster)) %>% 
  #generate cluster percentage for each school
  group_by(cluster, year, school_id) %>% 
  summarize(total = n_distinct(var),
            n = sum(usage),
            pct = n/total) %>% 
  select(school_id, year, cluster, n, pct) %>% 
  ungroup() %>% 
  pivot_wider(names_from = cluster,
              values_from = c(n, pct))
#read in characteristic vars for modeling and merge
vars <- import(here("data/longitudinal", "longitudinal_data.csv")) %>% 
  select(school_id, year, school_type, school_locale, school_enrollment, pct_bipoc, pct_ell, pct_frpl, pct_swd, grades_pk, grades_elementary, grades_middle, grades_high) %>% 
  mutate(c_enrollment = scale(school_enrollment, center = TRUE, scale = TRUE)[,1],
         c_bipoc = scale(pct_bipoc, center = TRUE, scale = TRUE)[,1],
         c_ell = scale(pct_ell, center = TRUE, scale = TRUE)[,1],
         c_frpl = scale(pct_frpl, center = TRUE, scale = TRUE)[,1],
         c_swd = scale(pct_swd, center = TRUE, scale = TRUE)[,1]) %>% 
  select(-c(school_enrollment, starts_with("pct")))
#merge
mod_dat <- mod_dat %>% 
  left_join(vars, by = c("school_id", "year"))
# set vars of interest
outcome <- mod_dat %>% select(starts_with("n_")) %>% colnames()
preds <- mod_dat %>% select(-school_id, -starts_with("pct_"), -starts_with("n_")) %>% colnames()
```

## Poisson

```{r}
#run models
model_list <- list()

for (outcome in outcome) {
  formula <- as.formula(paste(outcome, "~", paste(preds, collapse = "+")))
  model <- glm(formula, data = mod_dat, family = "poisson")
  model_list[[outcome]] <- model
}

#extract coeffs
tidy_models <- list()

for (outcome in names(model_list)) {
  tidy_model <- tidy(model_list[[outcome]])
  tidy_model$outcome <- outcome
  tidy_models[[outcome]] <- tidy_model
}

#combine
plot_data <- do.call(rbind, tidy_models) %>% 
  #rename outcome
  mutate(outcome = case_when(
    outcome == "n_deeper" ~ "Deeper learning",
    outcome == "n_ed_justice" ~ "Educational justice",
    outcome == "n_individualized" ~ "Individualized learning",
    outcome == "n_postsecondary" ~ "Postsecondary pathways",
    outcome == "n_misc" ~ "Misc. cluster"
  ),
  #rename terms
  term = str_remove_all(term, "school_type"),
  term = str_remove_all(term, "grades_"),
  term = str_remove_all(term, "c_"),
  term = str_remove_all(term, "school_locale")
  ) %>% 
  arrange(desc(estimate)) %>% 
  mutate(term = fct_inorder(term))
```

```{r}
#prep data
plot_data_list <- list()

for (outcome in names(model_list)) {
  tidy_model <- tidy(model_list[[outcome]], conf.int = TRUE)
  tidy_model$outcome <- outcome 
  plot_data_list[[outcome]] <- tidy_model
}

#create plot for each outcome
create_plot_for_outcome <- function(data, outcome) {
  ggplot(data, aes(y = term, x = estimate)) +
    geom_linerange(aes(xmin = estimate - std.error, 
                       xmax = estimate + std.error), 
                   color = "blue") +
    geom_point() +
    theme_minimal() +
    theme(panel.grid.major.y = element_blank()) +
    labs(
      x = "Estimated effect on cluster",
      y = "",
      title = sprintf("School characteristics describing %s", outcome)
    )
}

plots <- list()
for (outcome in names(plot_data_list)) {
  plots[[outcome]] <- create_plot_for_outcome(plot_data_list[[outcome]], outcome)
}
plots$n_deeper
plots$n_ed_justice
plots$n_individualized
plots$n_misc
plots$n_postsecondary
```

## Binomial logistic regression

```{r}
#add cluster n's for cbind()
mod_dat <- mod_dat %>% 
  mutate(tot_deeper = 12,
         tot_ed_justice = 19,
         tot_individualized = 9,
         tot_misc = 24,
         tot_postsecondary = 9)
# outcome_tot <- mod_dat %>% select(starts_with("tot_")) %>% colnames()
#manual run because loops aren't working
deep_mod <- glmer(cbind(n_deeper, tot_deeper - n_deeper) ~ year + school_type + school_locale + grades_pk + grades_elementary + grades_high + c_enrollment + c_bipoc + c_ell + c_frpl + c_swd + (1|school_id), 
                family = binomial,
                data = mod_dat)
# plot
tidy(model, effects = "ran_pars", conf.int = TRUE) %>% 
ggplot(., aes(y = term, x = estimate)) +
    geom_linerange(aes(xmin = estimate - std.error, 
                       xmax = estimate + std.error), 
                   color = "blue") +
    geom_point() +
    theme_minimal() +
    theme(panel.grid.major.y = element_blank()) +
    labs(
      x = "Estimated effect on cluster",
      y = "",
      title = sprintf("School characteristics describing %s", outcome)
    )
```

```{r, include = FALSE, eval = FALSE}
#run models
model_list <- list()

for (i in seq_along(outcome)) {
    out <- outcome[i]
    tot <- outcome_tot[i]

    formula_str <- paste("cbind(", out, ",", tot, "-", out, ")", "~", paste(preds, collapse = "+"), "+ (1|school_id)")
    formula <- as.formula(formula_str)
    model <- glmer(formula, data = mod_dat, family = binomial)
    model_list[[outcome]] <- model
}

#extract coeffs
tidy_models <- list()

for (outcome in names(model_list)) {
  tidy_model <- tidy(model_list[[outcome]])
  tidy_model$outcome <- outcome
  tidy_models[[outcome]] <- tidy_model
}

#combine
plot_data <- do.call(rbind, tidy_models) %>% 
  arrange(desc(estimate)) %>% 
  mutate(term = fct_inorder(term))
```
```

