---
title: "Preliminary Look at Schools over Time"
author: "Gregor Thomas"
date: "2023-12-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
pacman::p_load(tidyverse, here, rio, DT, ggcorrplot, forcats, lme4)
source(here("scripts/branding.R"))
#read in data
long = import(here("data/longitudinal", "tags-long.csv"))
wide = import(here("data/longitudinal", "tags-wide.csv"))
sch = import(here("data/long_school.csv")) |>
  summarize(n_years = n_distinct(year), .by = c(school_id, school_name))
#read in tag labels
#labels <- import(here("data", "tag_labels.csv"))
```

> The object of inquiry here is the school - how has each school's number of reported practices changed?

We'll begin by including schools that participated in 2, 3, and 4 surveys, and 
looking at how many practices they selected each year in total. (The blue line
shows the average, and the points for each school are conected by light gray lines.)


```{r, fig.width = 9, fig.height = 4}
school_summary = long |>
  filter(usage == 1) |>
  mutate(
    n_years = n_distinct(year),
    .by = school_id
  ) |>
  filter(n_years > 1) |>
  summarize(
    n_tags = sum(usage),
    .by = c(year, school_id, n_years)
  )

overall_summary = school_summary |>
  summarize(avg_tags = mean(n_tags), .by = c(year, n_years))

ggplot(school_summary, aes(x = year, y = n_tags)) +
  geom_line(aes(group = school_id), alpha = 0.2) +
  geom_point(aes(group = school_id), alpha = 0.2, color = "blue") +
  geom_line(data = overall_summary, aes(y = avg_tags), color = transcend_cols[1]) +
  facet_wrap(vars(paste(n_years, "years participation"))) +
  labs(
    title = "# of practices selected by schools (all practices)"
  ) +
  angle_text_x()
```

The above reflects the number of practices available to select, it is effected
by what practices schools indicate, but also the number of choices they were given
each year. Below, we recreate the same graph but limit it to practices that were
included in all 4 surveys. Comparing the two, we see above decreases in the averages,
but this is mostly due to culling of tags--below we can see that the averages
increase over time.

```{r, fig.width = 9, fig.height = 4}
## get vector of 4-year tags
tags_4y = long |>
  filter(usage == 1) |>
  filter(n_distinct(year) == 4, .by = var) |>
  pull(var)

long_4y = long |> filter(var %in% tags_4y)

school_summary_4p = long_4y |>
  mutate(
    n_years = n_distinct(year),
    .by = school_id
  ) |>
  filter(n_years > 1) |>
  summarize(
    n_tags = sum(usage),
    .by = c(year, school_id, n_years)
  )

overall_summary_4p = school_summary_4p |>
  summarize(avg_tags = mean(n_tags), .by = c(year, n_years))

ggplot(school_summary_4p, aes(x = year, y = n_tags)) +
  geom_line(aes(group = school_id), alpha = 0.2) +
  geom_point(aes(group = school_id), alpha = 0.2, color = "blue") +
  geom_line(data = overall_summary_4p, aes(y = avg_tags), color = transcend_cols[1]) +
  facet_wrap(vars(paste(n_years, "years participation"))) +
  labs(
    title = "# of practices selected by schools (4-year practices only)"
  ) +
  angle_text_x()
```

The table below shows the blue-line averages of the above graph (with 4-year practices).
There we can see that while the average number of selected practices by about 6-7
practices over the entire 2019-2023 period (focusing mostly on the schools responding 
in 3 or 4 years). It seems safe to say that repeat Canopy schools typically add 1-2 practices per year.


```{r}
overall_summary_4p |>
  arrange(n_years, year) |>
  datatable(
    rownames = FALSE,
    caption = "Average number of practices selected in each survey year, by the number of years the school responded to the survey (blue lines in the graph above). (Restricted to practices present on all 4 surveys.)",
    options = list(autoWidth = TRUE, iDisplayLength = 25)
  ) |>
  formatRound("avg_tags", digits = 1) |>
  formatStyle(
    'n_years', 
    backgroundColor = styleEqual(levels = c(2, 3, 4), values = transcend_cols[2:4])
  )
```

Which schools are adding and dropping practices? The table below shows, for
each school responding to multiple surveys, how many practices they added,
dropped, and waffled on (both added and dropped). While the average may be 6 or 7
practices added overall, plenty of schools have added 20+ practices.

I recommend digging deeper into the waffling, especially the few schools that have 
a lot of waffling. If a school indicates a practice was dropped and then re-added,
there are possibilities of data compilation error,  survey response error, or a COVID blip, 
and we may not want to read too much into any of those.

I'd also be interested in the schools that appear to have made drastic changes--if
those changes can be corroborated and explained with some interviews or even just
another survey response, it may be worth targeting them for some outreach.

```{r}
changes = long_4y |>
  arrange(school_id, var, year) |>
  summarize(
    added = any(diff(usage) == 1),
    dropped = any(diff(usage) == -1),
    waffled = added & dropped,
    .by = c(var, school_id)
  ) |>
  filter(added | dropped)

school_changes = changes |>
  summarize(
    across(c(added, dropped, waffled), sum),
    .by = school_id
  ) |>
  mutate(net_adds = added - dropped) |>
  arrange(desc(net_adds), desc(added)) |>
  left_join(sch, by = "school_id") |>
  mutate(school_id = fct_inorder(as.character(school_id))) |>
  select(school_id, school_name, n_years, added, dropped, waffled, net_adds) 

school_changes |>
  datatable(
    rownames = FALSE,
    caption = "Number of 4-survey-year practices added, dropped, and waffled (both added and dropped) by schools with multiple survey responses.") 
```

## Focus on the Wafflers

Purdue Polytechnic and Juab HS waffled a lot. Here's the raw data, including all
practices that they selected in at least 1 year.

```{r}
wafflers = c(105, 165)
waffle_detail = long_4y |>
  filter(school_id %in% wafflers) |>
  filter(all(0:1 %in% usage), .by = c(var, school_id)) |>
  mutate(
    added = any(diff(usage) == 1),
    dropped = any(diff(usage) == -1),
    waffled = added & dropped,
    .by = c(var, school_id)
  ) |>
  pivot_wider(names_from = year, values_from = usage) |>
  left_join(schools, by = "school_id") |>
  select(school_id, school_name, tag = var, `2019`:`2023`, added, dropped, waffled) |>
  arrange(school_id, desc(waffled))

waffle_detail |>
  datatable(rownames = FALSE)
```

Here's a summary of the patterns, where 1 indicates selected and 0 indicates
not selected in each survey, respectively. E.g., "1,0,1,1" means the practice was
selected in 2019, not selected in 2021, and then selected again in 2022 and 2023.
`n` is the number of practices for which this pattern was observed. We can see
that Juab HS has 11 tags with the `1,0,1,1` pattern, meaning there are 11 
practices that Juab indicated in all survey years *except* 2021.

```{r}
waffle_detail |>
  mutate(binary_summary = paste(`2019`, `2021`, `2022`, `2023`, sep = ",")) |>
  count(school_name, binary_summary) |>
  datatable(rownames = FALSE, options = list(autoWidth = TRUE, iDisplayLength = 25))
```


## Practices added and dropped

The next obvious question is which practices are schools adding and dropping?
Again, we look at the tags present on all 4 surveys and see how many multiply-responding
schools added or dropped them. The "waffled" column counts schools that both added
and dropped the practice

```{r}
tag_changes = changes |>
  summarize(
    across(c(added, dropped, waffled), sum),
    .by = var
  ) |>
  mutate(net_adds = added - dropped) |>
  arrange(net_adds, added) |>
  mutate(var = fct_inorder(var))

tag_changes |> arrange(desc(net_adds), desc(added)) |> datatable()
```

The same information from the table above is presented in the graph below. 
Unsurprisingly, far more practices were added that dropped, with only 4 practices
showing a net decrease. *Culturally Responsive* also is notable as the only practice
that was not dropped at all, only added.

I am surprised by how many drops there are - 10 practices are dropped by 20 or more schools,
and the 3 most dropped practices are dropped by 40+ schools. Those 3 are also
the highest on the "waffle" count and should be investigated more deeply before
we jump to conclusions. I can imagine a COVID-effect or an error in combining the
data.


```{r, fig.height = 8, fig.width = 10}
ggplot(tag_changes, aes(y = var)) +
  geom_segment(aes(x = dropped, xend = added, yend = var, color = factor(sign(net_adds)))) +
  geom_point(aes(x = dropped), color = "red") +
  geom_point(aes(x = added), color = "blue") +
  geom_text(
    aes(x = pmax(added, dropped), label = paste("Δ =", net_adds), color = factor(sign(net_adds))),
    nudge_x = 1,
    hjust = 0,
    show.legend = FALSE
  ) +
  scale_color_manual(
    values = c("red", "forestgreen", "blue"),
    labels = c("Drops", "Net 0", "Adds")
  ) +
  scale_y_discrete(labels = trimws) +
  guides(col = guide_legend(nrow = 1, title = NULL)) + 
  bar_x_scale_count +
  labs(
    y = "",
    x = "Number of schools adding/droping",
    title = "4-year tags added and dropped"
  ) +
  theme(
    panel.grid.major.y = element_blank(),
    legend.position = "bottom",
    axis.text.y = element_text(size = rel(0.6))
  ) 
```

## Broadening: Practices in 2+ years

Repeating the above analyses, but expanding the scope to include practices 
on the survey in 2023 and at least 1 other year.

```{r}
## find tags used in 2023 and at least 1 other year
tags_2y = long |>
  filter(usage == 1) |>
  filter(n_distinct(year) > 2 & 2023 %in% year, .by = var) |>
  pull(var)

long_2y = long |> filter(var %in% tags_2y)

changes_2y = long_2y |>
  arrange(school_id, var, year) |>
  summarize(
    added = any(diff(usage) == 1),
    dropped = any(diff(usage) == -1),
    waffled = added & dropped,
    .by = c(var, school_id)
  ) |>
  filter(added | dropped)

school_changes_2y = changes_2y |>
  summarize(
    across(c(added, dropped, waffled), sum),
    .by = school_id
  ) |>
  mutate(net_adds = added - dropped) |>
  arrange(desc(net_adds), desc(added)) |>
  left_join(sch, by = "school_id") |>
  mutate(school_id = fct_inorder(as.character(school_id))) |>
  select(school_id, school_name, n_years, added, dropped, waffled, net_adds) 

school_changes_2y |>
  datatable(
    rownames = FALSE,
    caption = "Number of practices from 2023 and at least one other year added, dropped, and waffled (both added and dropped) by schools with multiple survey responses.") 
```

Note that a practice included in only two years couldn't possibly be "waffled",
so there will be relatively fewer waffles for these 2-year tags, and even 3-year
tags have less opportunity to waffle.

```{r}
tag_changes_2y = changes_2y |>
  summarize(
    across(c(added, dropped, waffled), sum),
    .by = var
  ) |>
  mutate(net_adds = added - dropped) |>
  arrange(net_adds, added) |>
  mutate(var = fct_inorder(var))

tag_changes_2y |> arrange(desc(net_adds), desc(added)) |> datatable()
```

```{r, fig.height = 10, fig.width = 10}
ggplot(tag_changes_2y, aes(y = var)) +
  geom_segment(aes(x = dropped, xend = added, yend = var, color = factor(sign(net_adds)))) +
  geom_point(aes(x = dropped), color = "red") +
  geom_point(aes(x = added), color = "blue") +
  geom_text(
    aes(x = pmax(added, dropped), label = paste("Δ =", net_adds), color = factor(sign(net_adds))),
    nudge_x = 1,
    hjust = 0,
    show.legend = FALSE
  ) +
  scale_color_manual(
    values = c("red", "forestgreen", "blue"),
    labels = c("Drops", "Net 0", "Adds")
  ) +
  scale_y_discrete(labels = trimws) +
  guides(col = guide_legend(nrow = 1, title = NULL)) + 
  bar_x_scale_count +
  labs(
    y = "",
    x = "Number of schools adding/droping",
    title = "2+ year tags added and dropped"
  ) +
  theme(
    panel.grid.major.y = element_blank(),
    legend.position = "bottom",
    axis.text.y = element_text(size = rel(0.6))
  ) 
```

The *practices_services_learning* outlier needs some investigation, the table shows 81 adds and 57 drops, including
30 waffles. 


# Little bit of modeling

Here we'll build a model for how volatile/dynamic school's tagging
practices are.

We'll look **within schools**. For all schools with multiple responses, we'll make 
a data point out of each consecutive survey response, and for the 
tags that were used in both of those response years, we will look at
the **number of practices changed** (number added plus number dropped). We will then model count of practice changes using school
demography and ecology as predictors.

The time periods are important too, and we know that the pandemic had a 
large effect on many practices. To try to capture the relevant time
details without overcomplicating the model, we will classify each time
transition into one of four cases:

1. `to_covid` for transitions from 2019 to 2020
2. `from_covid` for transitions from 2020 to a later year
3. `post_covid` for transitions that start after 2020
4. `skip_covid` for transitions from 2019 to a post-2020 year

For now, we will not put the number of years in the transition as a numeric
variable, as that will be mostly dependent on the transition types, though 
with more years of data we may want to include that (and perhaps keep things
focused post-pandemic).

```{r tag_volatility_model, message = FALSE, warning = FALSE}
volatility_data = 
  wide |>
  filter(n() > 1, .by = school_id) |>
  arrange(school_id, year) |>
  mutate(
    i_seq = row_number(),
    transition = case_when(
      i_seq == 1 ~ NA_character_,
      lag(year) == 2022 ~ 'post_covid',
      year == 2021 ~ 'to_covid',
      lag(year) == 2021 ~ 'from_covid',
      year > 2021 & lag(year) == 2019 ~ 'skip_covid',
      .default = "PROBLEM"
    ),
    across(
      .cols = starts_with("practices"),
      \(x) c(0, abs(diff(x))) 
      # ^^^
      # take absolute difference from prev row
      # stick 0 on the front for first row
      # NAs (when tags weren't in use) will stay as NAs
    ),
    n_changes = rowSums(across(.cols = starts_with("practices")), na.rm = TRUE),
    .by = school_id
  ) |>
  filter(i_seq > 1) |> 
  select(!starts_with("practices"))

## get demographics
xx = read_csv(here("data/longitudinal/combined-dat-ip.csv"))
xx = xx |>
  mutate(
    stu_poc_pct = 1 - self_reported_race_white,
  ) |>
  select(
    school_id,
    year = panel_year, 
    total_enroll = self_reported_total_enrollment,
    stu_poc_pct,
    locale, 
    title_i_schoolwide,
    starts_with("grades")
  ) |>
  mutate(
    enroll_scaled = (total_enroll - mean(total_enroll, na.rm = TRUE)) / 
      sd(total_enroll, na.rm = TRUE),
    locale = relevel(factor(locale), ref = "Urban")
  )

mm = left_join(volatility_data, xx, by = c("school_id", "year"))

mm = mm |>
  group_by(school_id) |>
  fill(everything(), .direction = "downup") |>
  ungroup()

mm = mm |>
  mutate(transition = relevel(factor(transition), ref = "post_covid"))

missing_count =  mm |> 
  is.na() |> 
  colSums() |> 
  sort(decreasing = TRUE) |> 
  data.frame(n_missing = _)

datatable(missing_count)

vol_mod = lm(
  n_changes ~ transition + stu_poc_pct + enroll_scaled + locale +
     grades_elementary + grades_middle + grades_high,
  data = mm
)

vol_coef = broom::tidy(vol_mod) |>
  arrange(desc(estimate)) |>
  mutate(term = fct_inorder(term)) 

ggplot(vol_coef, aes(y = term, x = estimate)) +
  geom_linerange(aes(xmin = estimate - std.error, xmax = estimate + std.error)) +
  geom_point() +
  theme(panel.grid.major.y = element_blank()) +
  labs(
    x = "Estimated effect on practice changes",
    y = "",
    title = "Coefficient estimates for number of tags\nmodified by schools between survey years"
  )

```


```{r n_tags_model, include = FALSE, eval=FALSE}
## old code to model Number of tags each schools chose by year
## model didn't converge, and probably isn't that useful...
## get tags by year for the offset
n_tags_by_year = long |>
  summarize(n_sch = sum(usage), .by = c(var, year)) |>
  filter(n_sch > 0) |>
  summarize(n_tags_available = n(), .by = year)

## response and school ID
sch_total_tags = long |>
  summarize(n_tags = sum(usage), .by = c(school_id, year))

## get demographics
xx = read_csv(here("data/longitudinal/combined-dat-ip.csv"))
xx = xx |>
  mutate(
    stu_poc_pct = 1 - self_reported_race_white,
  ) |>
  select(
    school_id,
    year = panel_year, 
    total_enroll = self_reported_total_enrollment,
    stu_poc_pct,
    locale, 
    starts_with("grades")
  )

mod_dat = sch_total_tags |> left_join(xx, by = c("school_id", "year")) |>
  left_join(n_tags_by_year) |>
  mutate(year = factor(year)) |>
  mutate(total_enroll_std = scale(total_enroll))
  

mod = glmer(
  n_tags ~ year + total_enroll_std + stu_poc_pct + locale + 
    grades_elementary + grades_middle + grades_high + offset(log(n_tags_available)) + (1|school_id),
  data = mod_dat,
  family = poisson,
  )
```



```

```

