---
title: "Tag Frequency Changes"
author: "Janette Avelar"
date: '2023-12-16'
output: 
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warnings = FALSE, message = FALSE)
pacman::p_load(tidyverse, here, rio, DT, ggcorrplot, psych, parameters)
#read in data
dat <- import(here("data/longitudinal", "tags-long.csv"))
#read in tag labels
labels <- import(here("data", "tag_labels.csv"))
```

# Across schools

*Central Q: * For each practice where we have at least 2 years of data, how has the practice's frequency changed?  
-look only at changes within schools - so, schools that have at least 2 data points about that practice
-the object of inquiry here is the practice - how have each practice's frequencies changed?

A few ideas:  
-could be a table showing +/-X (where X is number of schools adding or removing) for each year. We did a graph version of this on p.8 in the 2021 report.  
-could be a line graph showing # of schools reporting each practice each year, similar to p.22 in the 2022 report. (Note, this code should be in our github from 2022.)  
-to make it more readable, it could be interactive (user could hover over a tag to see that line distinguished).  

### Raw Counts Table

```{r}
##### creating table with raw count differences across years #####
#starting with raw counts for each year
dat %>% 
  group_by(year, var) %>% 
  summarize(n = sum(usage, na.rm = TRUE)) %>% 
  #need to also figure out how many repeat over year
  ungroup() %>% 
  mutate(rate = rep(1, nrow(.))) %>% 
  group_by(var) %>% 
  mutate(rep = sum(rate, na.rm = TRUE)) %>% 
  ungroup() %>% 
  #drop vars without 2+ years of data
  filter(rep > 1) %>% 
  select(!c(rate, rep)) %>% 
  #make cols for each year to determine change
  pivot_wider(names_from = year,
              values_from = n) %>% 
  #calculate change
  mutate(change_2021 = `2021` - `2019`,
         change_2022 = `2022` - `2021`,
         change_2023 = `2023` - `2022`) %>% 
  #bring in labels
  left_join(labels, by = "var") %>% 
  select(Tag = label, `2019`, `2021` = change_2021, `2022` = change_2022, `2023` = change_2023) %>% 
  #table
  datatable() %>%
  #style
  formatStyle(columns = c("2021", "2022", "2023"), 
              color = styleInterval(cuts = 0, values = c("red", "green")),
              textAlign = "center")
```

### Percentage Table

```{r}
##### creating table with percentage differences across years #####
#year totals
totals <- dat %>% 
  select(school_id, year) %>% 
  unique() %>% 
  mutate(rate = rep(1, nrow(.))) %>% 
  group_by(year) %>% 
  summarize(n_school = sum(rate, na.rm = TRUE)) %>% 
  ungroup()
# create table with percentage changes
dat %>% 
  group_by(year, var) %>% 
  summarize(n = sum(usage, na.rm = TRUE)) %>% 
  left_join(totals, by = "year") %>% 
  mutate(pct = n/n_school,
         rate = rep(1, n())) %>% 
  ungroup() %>% 
  #need to also figure out how many repeat over year
  group_by(var) %>% 
  mutate(rep = sum(rate, na.rm = TRUE)) %>% 
  ungroup() %>% 
  #drop vars without 2+ years of data
  filter(rep > 1) %>% 
  select(var, year, pct) %>% 
  #make cols for each year
  pivot_wider(names_from = year,
              values_from = pct) %>%
  #calculate change
  mutate(change_2021 = `2021` - `2019`,
         change_2022 = `2022` - `2021`,
         change_2023 = `2023` - `2022`) %>% 
  #bring in labels
  left_join(labels, by = "var") %>% 
  select(Tag = label, `2019`, `2021` = change_2021, `2022` = change_2022, `2023` = change_2023) %>% 
  #table
  datatable() %>%
  #style
  formatStyle(columns = c("2021", "2022", "2023"), 
              color = styleInterval(cuts = 0, values = c("red", "green")),
              textAlign = "center") %>% 
  formatPercentage(c("2019", "2021", "2022", "2023"), 0)
```



# Within Schools

*Central Q: * For each practice where we have at least 2 years of data, how has the practice's frequency changed?  
-look only at changes within schools - so, schools that have at least 2 data points about that practice
-the object of inquiry here is the practice - how have each practice's frequencies changed?

A few ideas:  
-could be a table showing +/-X (where X is number of schools adding or removing) for each year. We did a graph version of this on p.8 in the 2021 report.  
-could be a line graph showing # of schools reporting each practice each year, similar to p.22 in the 2022 report. (Note, this code should be in our github from 2022.)  
-to make it more readable, it could be interactive (user could hover over a tag to see that line distinguished).  

### Raw Counts Table - 2 year tags

This table displays the raw count differences across years for tags that appear in 2 years of data.


```{r}
##### creating table with raw count differences across years #####
#starting with raw counts for each year
n_dat <- dat %>% 
  mutate(n_year_tag = n_distinct(year), .by = var) %>% 
  mutate(n_year_schl = n_distinct(year), .by = school_id)

# 2 year tags
#overall view table
overall_2yr <-
n_dat %>% 
  filter(n_year_tag == 2) %>% 
  # mutate(n = sum(usage), .by = c(year, var)) %>% 
  filter(n_year_schl > 2) %>% 
  select(-n_year_tag, -n_year_schl) %>% 
  # group_by(year, var) %>% 
  # summarize(n = sum(usage, na.rm = TRUE)) %>% 
  #make cols for each year to determine change
  pivot_wider(names_from = year,
              values_from = usage) %>% 
  #calculate change
  group_by(school_id, var) %>% 
  mutate(change_2021 = `2021` - `2019`,
         change_2022 = `2022` - `2021`,
         change_2023 = `2023` - `2022`,
         added_2021 = ifelse(change_2021 == 1, 1, 0),
         added_2022 = ifelse(change_2022 == 1, 1, 0),
         added_2023 = ifelse(change_2023 == 1, 1, 0),
         drop_2021 = ifelse(change_2021 == -1, 1, 0),
         drop_2022 = ifelse(change_2022 == -1, 1, 0),
         drop_2023 = ifelse(change_2023 == -1, 1, 0)) %>% 
  ungroup() %>% 
  group_by(var) %>% 
  summarize(`2019` = sum(`2019`, na.rm = TRUE),
            `2021` = sum(change_2021, na.rm = TRUE),
            `2022` = sum(change_2022, na.rm = TRUE),
            `2023` = sum(change_2023, na.rm = TRUE),
            `2021 adds` = sum(added_2021, na.rm = TRUE),
            `2022 adds` = sum(added_2022, na.rm = TRUE),
            `2023 adds` = sum(added_2023, na.rm = TRUE),
            `2021 drops` = sum(drop_2021, na.rm = TRUE),
            `2022 drops` = sum(drop_2022, na.rm = TRUE),
            `2023 drops` = sum(drop_2023, na.rm = TRUE)) %>% 
  mutate(across(`2019`:`2023 drops`, ~na_if(., 0))) %>% 
  #bring in labels
  left_join(labels, by = "var")

#overview table
overall_2yr %>% 
  select(Tag = label, `2019`, `2021`, `2022`, `2023`) %>% 
  #table
  datatable() %>%
  #style
  formatStyle(columns = c("2021", "2022", "2023"), 
              color = styleInterval(cuts = 0, values = c("red", "green")),
              textAlign = "center")
```

This table breaks down the raw count for 2-year tags, by displaying the number of adds and drops in each year.


```{r}
#adds and drops table
overall_2yr %>% 
  select(Tag = label, ends_with("adds"), ends_with("drops")) %>% 
  mutate(across(ends_with("drops"), ~.*-1)) %>% 
  mutate(across(`2021 adds`:`2023 drops`, ~na_if(., 0))) %>% 
  #table
  datatable() %>%
  #style
  formatStyle(columns = c("2021 adds", "2022 adds", "2023 adds", "2021 drops", "2022 drops", "2023 drops"), 
              color = styleInterval(cuts = 0, values = c("red", "green")),
              textAlign = "center")
```

### Raw Counts Table - 3 year tags

This table displays the raw count differences across years for tags that appear in 3 years of data.

```{r}
# 3 year tags
#overall view table
overall_3yr <-
n_dat %>% 
  filter(n_year_tag == 3) %>% 
  # mutate(n = sum(usage), .by = c(year, var)) %>% 
  filter(n_year_schl > 3) %>% 
  select(-n_year_tag, -n_year_schl) %>% 
  # group_by(year, var) %>% 
  # summarize(n = sum(usage, na.rm = TRUE)) %>% 
  #make cols for each year to determine change
  pivot_wider(names_from = year,
              values_from = usage) %>% 
  #calculate change
  group_by(school_id, var) %>% 
  mutate(change_2021 = `2021` - `2019`,
         change_2022 = `2022` - `2021`,
         change_2023 = `2023` - `2022`,
         added_2021 = ifelse(change_2021 == 1, 1, 0),
         added_2022 = ifelse(change_2022 == 1, 1, 0),
         added_2023 = ifelse(change_2023 == 1, 1, 0),
         drop_2021 = ifelse(change_2021 == -1, 1, 0),
         drop_2022 = ifelse(change_2022 == -1, 1, 0),
         drop_2023 = ifelse(change_2023 == -1, 1, 0)) %>% 
  ungroup() %>% 
  group_by(var) %>% 
  summarize(`2019` = sum(`2019`, na.rm = TRUE),
            `2021` = sum(change_2021, na.rm = TRUE),
            `2022` = sum(change_2022, na.rm = TRUE),
            `2023` = sum(change_2023, na.rm = TRUE),
            `2021 adds` = sum(added_2021, na.rm = TRUE),
            `2022 adds` = sum(added_2022, na.rm = TRUE),
            `2023 adds` = sum(added_2023, na.rm = TRUE),
            `2021 drops` = sum(drop_2021, na.rm = TRUE),
            `2022 drops` = sum(drop_2022, na.rm = TRUE),
            `2023 drops` = sum(drop_2023, na.rm = TRUE)) %>% 
  mutate(across(`2019`:`2023 drops`, ~na_if(., 0))) %>% 
  #bring in labels
  left_join(labels, by = "var")

#overview table
overall_3yr %>% 
  select(Tag = label, `2019`, `2021`, `2022`, `2023`) %>% 
  #table
  datatable() %>%
  #style
  formatStyle(columns = c("2021", "2022", "2023"), 
              color = styleInterval(cuts = 0, values = c("red", "green")),
              textAlign = "center")
```

This table breaks down the raw count for 3-year tags, by displaying the number of adds and drops in each year.

```{r}
#adds and drops table
overall_3yr %>% 
  select(Tag = label, ends_with("adds"), ends_with("drops")) %>% 
  mutate(across(ends_with("drops"), ~.*-1)) %>% 
  mutate(across(`2021 adds`:`2023 drops`, ~na_if(., 0))) %>% 
  #table
  datatable() %>%
  #style
  formatStyle(columns = c("2021 adds", "2022 adds", "2023 adds", "2021 drops", "2022 drops", "2023 drops"), 
              color = styleInterval(cuts = 0, values = c("red", "green")),
              textAlign = "center")
```

### Raw Counts Table - 4 year tags

This table displays the raw count differences across years for tags that appear in all 4 years of data.

```{r}
# 4 year tags
#overall view table
overall_4yr <-
n_dat %>% 
  filter(n_year_tag == 4) %>% 
  # mutate(n = sum(usage), .by = c(year, var)) %>% 
  filter(n_year_schl == 4) %>% 
  select(-n_year_tag, -n_year_schl) %>% 
  # group_by(year, var) %>% 
  # summarize(n = sum(usage, na.rm = TRUE)) %>% 
  #make cols for each year to determine change
  pivot_wider(names_from = year,
              values_from = usage) %>% 
  #calculate change
  group_by(school_id, var) %>% 
  mutate(change_2021 = `2021` - `2019`,
         change_2022 = `2022` - `2021`,
         change_2023 = `2023` - `2022`,
         added_2021 = ifelse(change_2021 == 1, 1, 0),
         added_2022 = ifelse(change_2022 == 1, 1, 0),
         added_2023 = ifelse(change_2023 == 1, 1, 0),
         drop_2021 = ifelse(change_2021 == -1, 1, 0),
         drop_2022 = ifelse(change_2022 == -1, 1, 0),
         drop_2023 = ifelse(change_2023 == -1, 1, 0)) %>% 
  ungroup() %>% 
  group_by(var) %>% 
  summarize(`2019` = sum(`2019`, na.rm = TRUE),
            `2021` = sum(change_2021, na.rm = TRUE),
            `2022` = sum(change_2022, na.rm = TRUE),
            `2023` = sum(change_2023, na.rm = TRUE),
            `2021 adds` = sum(added_2021, na.rm = TRUE),
            `2022 adds` = sum(added_2022, na.rm = TRUE),
            `2023 adds` = sum(added_2023, na.rm = TRUE),
            `2021 drops` = sum(drop_2021, na.rm = TRUE),
            `2022 drops` = sum(drop_2022, na.rm = TRUE),
            `2023 drops` = sum(drop_2023, na.rm = TRUE)) %>% 
  mutate(across(`2019`:`2023 drops`, ~na_if(., 0))) %>% 
  #bring in labels
  left_join(labels, by = "var")

#overview table
overall_2yr %>% 
  select(Tag = label, `2019`, `2021`, `2022`, `2023`) %>% 
  #table
  datatable() %>%
  #style
  formatStyle(columns = c("2021", "2022", "2023"), 
              color = styleInterval(cuts = 0, values = c("red", "green")),
              textAlign = "center")
```

This table breaks down the raw count for 4-year tags, by displaying the number of adds and drops in each year.

```{r}
#adds and drops table
overall_2yr %>% 
  select(Tag = label, ends_with("adds"), ends_with("drops")) %>% 
  mutate(across(ends_with("drops"), ~.*-1)) %>% 
  mutate(across(`2021 adds`:`2023 drops`, ~na_if(., 0))) %>% 
  #table
  datatable() %>%
  #style
  formatStyle(columns = c("2021 adds", "2022 adds", "2023 adds", "2021 drops", "2022 drops", "2023 drops"), 
              color = styleInterval(cuts = 0, values = c("red", "green")),
              textAlign = "center")
```

